### 2019-12-03 annotations/face_hands/analysis/analyze_goldset.Rmd
Add relevant csvs for turk hits, start gold set annotation analysis

---
title: "Face/Hand Annotations "
output: html_notebook
---

```{r setup, echo = FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(lmerTest)
library(viridis)

library(egg)
theme_set(theme_few())
```

```{r message=FALSE}
d.raw <- data.frame()
trial_indexes = seq(1,252,1)

files <- dir("turk_annotations/")
for (f in files) {
  jf <- paste("turk_annotations/",f,sep="")
  jd <- read_csv(jf)

  ## really not sure why there's this weird discrepancy for the last trial, but OK
  jd$`Answer.trial-252` = jd$`Answer.trial-252.label`

  for (trial_ind in trial_indexes){
    trial_data = data.frame(
      ## including HITId and HITTYpe
      this_response = eval(parse(text = paste0("jd$`Answer.trial-",trial_ind,"`"))),
      full_image_path = eval(parse(text = paste0("jd$Input.img_src",trial_ind))),
      trial_number = trial_ind,
      HITID = jd$HITId,
      HITTypeId =jd$HITTypeId
    )
    d.raw <- bind_rows(d.raw, trial_data)
  }
}

```