### 2019-12-03 annotations/face_hands/analysis/analyze_goldset.Rmd
Update annotations preprocessing script

---
title: "Face/Hand Annotations "
output: html_notebook
---

```{r setup, echo = FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(lmerTest)
library(viridis)

library(egg)
theme_set(theme_few())
```

```{r message=FALSE}
d.raw <- data.frame()
trial_indexes = seq(1,252,1)

## function to spit back category of sanity check images
sanity_check_category = function(shorter_image_path){
 ## I hate this function but I couldn't make it work without 
## going down a regex rabbithole so here it is
  hands = str_detect(shorter_image_path,'hands')
  faces = str_detect(shorter_image_path,'faces')
  none = str_detect(shorter_image_path,'none')
  both = str_detect(shorter_image_path,'face-hands')

 if (both==TRUE){
   category = 'both'
 }
 else if (hands==TRUE){
   category = 'hands'
 }
 else if (faces==TRUE){
   category='faces'
 }
 else if (none==TRUE){
  category='none'   
 }
 else{
   category=NA
  }

 return(category)
}

```

```{r}
files <- dir("turk_annotations/")
for (f in files) {
  jf <- paste("turk_annotations/",f,sep="")
  jd <- read_csv(jf)

  ## really not sure why there's this weird discrepancy for the last trial, but OK
  jd$`Answer.trial-252` = jd$`Answer.trial-252.label`

  for (trial_ind in trial_indexes){
    trial_data = data.frame(
      ## including HITId and HITTYpe
      this_response = eval(parse(text = paste0("jd$`Answer.trial-",trial_ind,"`"))),
      full_image_path = eval(parse(text = paste0("jd$Input.img_src",trial_ind))),
      trial_number = trial_ind,
      HITID = jd$HITId,
      HITTypeId =jd$HITTypeId
    )
    d.raw <- bind_rows(d.raw, trial_data)
  }
}

### Pop out sanity check category from image path
d.raw <- d.raw %>%
    mutate(short_image_path = str_split_fixed(full_image_path,'/',7)[,7]) %>%
    mutate(shorter_image_path = str_split_fixed(short_image_path,'.jpg',2)[,1]) %>%
    mutate(sanity_check_category = sanity_check_category(shorter_image_path))

### Recode responses into easier to munge format
level_key <- c("Face(s) and Hand(s)" = "both", "Hand(s) only" = "hands", "Neither face(s) nor hands(s)" = "none", "Face(s) only" = "faces")

d.raw$response=array()
d.raw$response <- recode(d.raw$this_response, !!!level_key)

### Break out sanity check images and look at discrepancies
sanity_images <- d.raw %>%
  filter(!is.na(sanity_check_category)) %>%
  mutate(correct_sanity = response == sanity_check_category)

```

```{r}

```