### 2019-12-05 analysis/analysis_dec2019.Rmd
Merging detections with goldset, rendering examples

---
title: "Analysis_dec2019"
author: "Bria Long"
date: "12/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(egg)
theme_set(theme_few())
```

## Read in list of detections (takes forever)
```{r cars}
data_file = "../data/all_frames.csv"
detections =read.csv(data_file)
```

### Filter flipped videos
```{r}
right_side_up_file = "video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

# get INCORRECT videos
flipped_videos <- right_side_up %>%
  filter(right_side_up==1)

d <- detections %>%
  filter(!vid_name %in% flipped_videos$video)
```

## Read in turk annotations
```{r}
## Need to be able to merge turk annotations with openpose detections but don't understand how vid_name and frame number map onto .jpg names in the image names -- assuming sample index is first b/c max val ~24K

turk_annotations = "../data/turk_annotations_dec5.csv" # first 20 turkers
gold_set = read.csv(turk_annotations)

## cleanup goldset annotations so we can merge
g <- gold_set %>%
  filter(is.na(sanity_check_category)) %>% # don't consider sanity check images here
  filter(!is.na(hands)) %>%
  filter(!is.na(faces)) %>%
  mutate(sample_ind = str_split_fixed(short_image_path,"-",3)[,1], frame_ind_temp = str_split_fixed(short_image_path,"-",3)[,3], vid_name_temp = str_split_fixed(short_image_path,"-",3)[,2]) %>%
  mutate(sample_ind = as.numeric(sample_ind), frame_ind = as.numeric(str_split_fixed(frame_ind_temp,".jpg",3)[,1])) %>%
  filter(!is.na(sample_ind))  %>%
  mutate(vid_name = str_split(vid_name_temp,".mp4")[[1]][1]) %>%
  filter(!vid_name %in% flipped_videos$video)

## standardize for merging with detectinos
g$vid_name = as.factor(g$vid_name)
g$frame = g$frame_ind

# what <- gold_set %>%
#   filter(is.na(sample_ind))
```

```{r}
# no_openpose <- merged %>%
#   filter(is.na(face_openpose))
# 
# ## never run on openpose?
# unique(no_openpose$vid_name)
# "A_20150507_3124_01"

merged <- g %>%
  left_join(detections)  %>%
  filter(!is.na(face_openpose))
```

```{r}
# Function to evaluate detectors
evaluate_detector <- function(truth, detection) {
  if (truth == TRUE) {
    if (truth == detection) return ("TP") # was face/wrist, detected face/wrist
    else return("FN") # was face/wrist, missed face/wrist
  }
  else {
    if (truth == detection) return("TN") # was not face/wrist, did not detect face/wrist
    else return("FP") # was not face/wrist, detected face/wrist
  }
}

m <- merged %>%
  select(vid_name, frame, response, hands, faces, sample_ind, face_openpose, hand_openpose, short_image_path, full_image_path ) %>%
  mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
  rowwise() %>%
  mutate(face_eval = evaluate_detector(faces, face_openpose), hand_eval = evaluate_detector(hands, hand_openpose))
```

### Now calculate prfs - these are SUPER bad and not sure what is up.
```{r}
 return_prf = function(eval){
  tp=sum(eval == "TP")
  fp=sum(eval == "FP")
  fn=sum(eval == "FN")
  p = tp / (tp + fp)
  r = tp / (tp + fn)
  f=( 2 * p * r )/ (p + r)
  return(c(p,r,f))
}

return_prf(m$face_eval)
return_prf(m$hand_eval)

```

### Let's sanity check -- only look at face images from gold set...
```{r}
faces_only <- m %>%
  filter(faces==TRUE)

# yep these are correct
unique(faces_only$response)

# let's look at the actual iamges
dir.create(paste0('faces_check/'))
for (i in seq(1,length(faces_only$full_image_path),1)){
  image_read(as.character(faces_only$full_image_path[i])) %>%
  image_append(stack = FALSE) %>%
  image_write(file.path(paste0("faces_check/", faces_only$short_image_path[i])))
}
```

## how about true positives on faces?
```{r}
faces_only_op_cor <- m %>%
  filter(faces == TRUE) %>%
  filter(face_openpose == TRUE)

# how about images where open pose got it? these should be most obvious
dir.create(paste0('faces_tp/'))
for (i in seq(1,length(faces_only_op_cor$full_image_path),1)){
  image_read(as.character(faces_only_op_cor$full_image_path[i])) %>%
  image_append(stack = FALSE) %>%
  image_write(file.path(paste0("faces_tp/", faces_only_op_cor$short_image_path[i])))
}

```

## how about false positives on facs?
```{r}
to_write <- m %>%
  filter(faces == FALSE) %>%
  filter(face_openpose == TRUE)

# how about images where open pose got it? these should be most obvious
dir.create(paste0('faces_fp/'))
for (i in seq(1,length(to_write$full_image_path),1)){
  image_read(as.character(to_write$full_image_path[i])) %>%
  image_append(stack = FALSE) %>%
  image_write(file.path(paste0("faces_fp/", to_write$short_image_path[i])))
}

```

## how about false negatives on faces?
```{r}
to_write <- m %>%
  filter(faces == TRUE) %>%
  filter(face_openpose == FALSE)

# how about images where open pose got it? these should be most obvious
dir.create(paste0('faces_fn/'))
for (i in seq(1,length(to_write$full_image_path),1)){
  image_read(as.character(to_write$full_image_path[i])) %>%
  image_append(stack = FALSE) %>%
  image_write(file.path(paste0("faces_fn/", to_write$short_image_path[i])))
}

```

# Examine age-trends in open-pose detections
### Make age bins
```{r}
### What ages do we have detections at?

hist(d$age_days)
## bin age by 10 days
bin_size = 30
min_age = min(d$age_days)
max_age = max(d$age_days)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)
```

### Make detections into logical vectors and compute "person" as either face/hand
```{r}
d$face_openpose = as.logical(d$face_openpose)
d$hand_openpose = as.logical(d$hand_openpose)
d$person_openpose = (as.logical(d$face_openpose) | as.logical(d$hand_openpose))
d$both_openpose = (as.logical(d$face_openpose) & as.logical(d$hand_openpose))

```

```{r}
summary_by_age <- d %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  group_by(age_day_bin, child_id) %>%
  summarize(num_detect = length(face_openpose), prop_faces = sum(face_openpose) / num_detect,  prop_hands = sum(hand_openpose) / num_detect, face_or_hand = sum(person_openpose)/ num_detect, face_and_hand = sum(both_openpose)/ num_detect)

# for geom_smooth
summary_by_age$age_day_bin=as.numeric(summary_by_age$age_day_bin)
```

### Make diagnotic plots
```{r}
# person by age
ggplot(summary_by_age, aes(x=age_day_bin, y=face_or_hand, size=num_detect, col=child_id)) +
  geom_point(alpha=.8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_smooth() + 
  ylab('Proportion Face OR Hand') + 
  xlab('Age bin')

# person by age
ggplot(summary_by_age, aes(x=age_day_bin, y=face_and_hand, size=num_detect, col=child_id)) +
  geom_point(alpha=.8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_smooth() + 
  ylab('Proportion Face AND Hand') + 
  xlab('Age bin')

# faces
p1 = ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces, size=num_detect, col=child_id)) +
  geom_point(alpha=.8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_smooth() + 
  ylab('Proportion Faces') + 
  xlab('Age bin')

# hands
p2 = ggplot(summary_by_age, aes(x=age_day_bin, y=prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_smooth() + 
  ylab('Proportion Hands') + 
  xlab('Age bin')
```

### Key plot for fausey comparison
```{r}
## faces - hands
p3 = ggplot(summary_by_age, aes(x=age_day_bin, y=prop_faces - prop_hands, size=num_detect, col=child_id)) +
  geom_point(alpha=.8) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  geom_smooth() + 
  ylab('Proportion Faces - Hands') + 
  xlab('Age bin')

```