### 2020-01-11 annotations/face_hands/analysis/preprocess_turk_segmentations.Rmd
preprocessing turk segmentations

---
title: "Preprocess turk annotations"
output: html_notebook
---

```{r setup, echo = FALSE}
library(knitr)
opts_chunk$set(echo = TRUE)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(lmerTest)
library(viridis)
library(magick)
library(egg)
library(jsonlite)
theme_set(theme_few())
```

# Preprocessing

```{r}
# f <- dir("turk_annotations_segmentation/")
# jf <- paste("turk_annotations_segmentation/",f,sep="")
# jd <- read_csv(jf)

f = "turk_segmentations_hands_only.csv"
jf <- paste("../../../data/",f,sep="")
jd <- read_csv(jf)

d.raw <- data.frame()
trial_indexes = seq(1,length(jd$Input.img_src1),1)
  for (trial_ind in trial_indexes){
    ## get annotation and see if empty
    this_annotation = jd$Answer.annotatedResult.boundingBoxes[trial_ind]
    cleaned_annotation = fromJSON(this_annotation)

    # if there was an annotation
    if (length(cleaned_annotation)>1){

        ## for each thing that was annotated, new row
        for (annotation_ind in seq(1,length(cleaned_annotation$label),1)){
              trial_data = data.frame(
              label = cleaned_annotation$label[annotation_ind],
              height = cleaned_annotation$height[annotation_ind],
              width = cleaned_annotation$width[annotation_ind],
              left = cleaned_annotation$left[annotation_ind],
              top = cleaned_annotation$top[annotation_ind],
              full_image_path = jd$Input.img_src1[trial_ind],
              trial_number = trial_ind,
              HITID = jd$HITId[trial_ind],
              HITTypeId =jd$HITTypeId[trial_ind],
              WorkerID = jd$WorkerId[trial_ind]
              )
        d.raw <- bind_rows(d.raw, trial_data)
        } # for annotation_ind
    } 
    ## if nothing was annotated
    else{
    trial_data = data.frame(
          ## including HITId and HITTYpe
          label = 'None',
          height = NA,
          width = NA,
          left = NA,
          top = NA,
          full_image_path = jd$Input.img_src1[trial_ind],
          trial_number = trial_ind,
          HITID = jd$HITId[trial_ind],
          HITTypeId =jd$HITTypeId[trial_ind],
          WorkerID = jd$WorkerId[trial_ind]
          )
     d.raw <- bind_rows(d.raw, trial_data)
    }
}
```

## look at distribution of images done per worker
```{r}
worker_summary <- jd %>%
  group_by(WorkerId) %>%
  summarize(count_images = n(), avg_time = mean(WorkTimeInSeconds))
```

```{r}
total_area = jd$Answer.annotatedResult.inputImageProperties.height[1] * jd$Answer.annotatedResult.inputImageProperties.width[1]
d <- d.raw %>%
  mutate(detection_area = (height*width)/total_area)

## plot detection area by detection type
ggplot(d, aes(x=label, y=log(detection_area))) +
  geom_boxplot()

## faces tend to appear towards the top of the image (i.e. 0 pixel coordinates, or upper visual field)
ggplot(d, aes(x=label, y=top)) +
  geom_boxplot()

```

```{r}

d <- d.raw %>%
  mutate(hands_seg = (label %in% c("Child hand","Adult hand")), face_seg = (label == "Adult Face")) %>%
  mutate(short_image_path = str_split_fixed(full_image_path,'/',7)[,7]) %>%
  mutate(sample_ind = str_split_fixed(short_image_path,"-",3)[,1], frame_ind_temp = str_split_fixed(short_image_path,"-",3)[,3], vid_name_temp = str_split_fixed(short_image_path,"-",3)[,2]) %>%
  mutate(sample_ind = as.numeric(sample_ind), frame_ind = as.numeric(str_split_fixed(frame_ind_temp,".jpg",3)[,1])) %>%
  rowwise() %>%
  mutate(vid_name_short = str_split(vid_name_temp,"[.]")[[1]][1]) %>%
  mutate(vid_name = str_split_fixed(vid_name_temp,".mp4",2)[,1]) 
```

### How many frames/videos had children's faces?
```{r}
child_face_frames <- d %>%
  filter(label=="Child face")

# render out some of these...
# dir.create(paste0('child_face/'))
# for (i in seq(1,length(child_face_frames$full_image_path),1)){
#   image_read(as.character(child_face_frames$full_image_path[i])) %>%
#   image_append(stack = FALSE) %>%
#   image_write(file.path(paste0("child_face/", child_face_frames$short_image_path[i])))
# }

# dir.create(paste0('annotated-images-child-face/'))
# unique_images = child_face_frames$full_image_path
# for (image in unique_images){
#   this_image_rgb <- image_read(image)
#   img <- image_draw(this_image_rgb)
#   a <- d %>%
#     filter(full_image_path %in% image) 
#   
### rect(xleft, ybottom, xright, ytop
#   rect(a$left, a$top+a$height, a$left+a$width, a$top)
#   text(a$left+10, a$top+10, a$label)
#   
#   annotated_file = paste0('annotated-images-child-face/',unique(a$short_image_path), 'turk-annotated.jpg')
#   image_write(img, annotated_file)
#   dev.off()
# }

```

### Write out csv for use in analysis
```{r}
write_csv(d,'turk_segmentations_hands_only_processed.csv')
```

```{r}
d <- d %>%
  mutate(short_image_path = str_split_fixed(full_image_path,'/',7)[,7])
```

```{r eval=FALSE}

unique_images = d$full_image_path
for (image in unique_images){
  this_image_rgb <- image_read(image)
  img <- image_draw(this_image_rgb)
  a <- d %>%
    filter(full_image_path %in% image)

  rect(a$left, a$top+a$height, a$left+a$width, a$top)
  text(a$left+10, a$top+10, a$label)

  annotated_file = paste0('annotated-images/',unique(a$short_image_path), 'turk-annotated.jpg')
  image_write(img, annotated_file)
  dev.off()
}

```

```{r}
faces <- d %>%
  filter(label =="Face")

rect(faces$left, faces$top+faces$height, faces$left+faces$width, faces$top)
text(faces$left+10, faces$top+10, faces$label)
```

```{r}  
child_hand <- d %>%
  filter(label =="Child hand")

rect(child_hand$left, child_hand$top+child_hand$height, child_hand$left+child_hand$width, child_hand$top)
text(child_hand$left+10, child_hand$top+10, child_hand$label)

```