### 2020-01-11 analysis/analysis_goldset.Rmd

---
title: "saycam goldset analysis"
author: "Bria Long"
date: "12/4/2019, continued 1/9/2020"
output: html_document
---

# Basic data loading and setup
## Packages
```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
theme_set(theme_few())
```

## Read in data 
### Load ketan's gold set annotations from preprocessed Rdata files
```{r}

### Get video/frame info from full cleaned dataset (saved)
# load('../data/openpose_detections_filtered.RData')
# vid_info <- d_all %>%
#   select(vid_name, frame, age_days, child_id)
# save(vid_info, file = '../data/vid_info.RData')

load('../data/vid_info.RData')
load('../data/ketan_gold_sample.RData')
```

### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
```{r}
right_side_up_file = "../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

# get videos that are flipped, i.e., 
flipped_videos <- right_side_up %>%
  filter(right_side_up==1) %>%
  rename(vid_name = video)
```

### Load SAYCAM metadata from csv file; get metadata from detections .csv
```{r}
metadata_file = "../data/SAYcam-2019-12-0922_54_28.csv"
meta=read.csv(metadata_file) %>%
  rowwise() %>%
  mutate(vid_name = str_split(File.Name,"[.]")[[1]][1]) %>%
  mutate(count_locations = length(str_split(Location,',')[[1]]))

# collapse duplicate labels
meta$Location[meta$Location=='Bedroom Alice/Sam'] = 'Bedroom'
```

### Import bounding boxes for goldset & preprocess: filter videos, gather metadata, convert to pixel coordinates
```{r}
bounding_boxes = "../data/gold_sample_bounding_boxes.csv" 
height_px = 480; width_px = 640

bb = read.csv(bounding_boxes) %>%
  rowwise() %>%
  mutate(vid_name = str_split(vid_name,"[.]")[[1]][1]) %>%
  filter(!vid_name %in% flipped_videos$vid_name) %>%
  left_join(vid_info) %>%
  mutate(height_y = height*height_px, width_x = width*width_px, area = (height_y*width_x)/(height_px*width_px)) 
```
### Read in turk segmentations to get child vs. adult hands
```{r}
turk_annotations_seg = "../data/turk_segmentations_hands_only_processed.csv" 
g_seg = read.csv(turk_annotations_seg) %>%
  select(-HITID, -HITTypeId, -WorkerID)

## preprocess such that each frame has a true/false value for whether there was a child or adult hand
child_adult_hand_annotations <- g_seg %>%
    mutate(child_hands = (label=="Child hand"), adult_hands = (label=="Adult hand")) %>%
    select(-vid_name) %>%
    rename(vid_name = vid_name_short) %>%
    group_by(full_image_path, vid_name, frame_ind) %>%
    summarize(child_hand_seg = (sum(child_hands))>0,  adult_hand_seg = (sum(adult_hands)>0)) %>%
    mutate(frame = frame_ind) %>%
    select(child_hand_seg, adult_hand_seg, full_image_path, vid_name, frame) 
```

### Preprocess gold set data
```{r}
 ketan_gold_out <- ketan_gold %>%
   rowwise() %>%
   mutate(vid_name = str_split(vid_name,"[.]")[[1]][1]) %>%
   mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
   rename(face_present_ketan = face_present, hand_present_ketan = hand_present)

### Join metadata with gold set annotations (hereafter 'd')
d <- ketan_gold_out %>%
  filter(!vid_name %in% flipped_videos$vid_name) %>%
  left_join(vid_info, by=c("vid_name","frame")) %>%
  left_join(meta, by="vid_name")
```

# Analyze gold set annotations
### Make bins for collapsing across anotations
```{r}
bin_size = 7 # so it's weeks or months
min_age = min(d$age_days)
max_age = max(d$age_days)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)
d <- d %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1)))

# and make numeric
d$age_day_bin <- as.numeric(as.character(d$age_day_bin))

```

### Hands vs. faces in manual annotations, broken down by child vs. adult hands
```{r}
face_vs_hands_goldset <- d %>%
  full_join(child_adult_hand_annotations, by=c("vid_name", "frame")) %>%
  mutate(frame_id = paste0(vid_name, '-', frame)) %>%
  replace_na(list(adult_hand_seg = FALSE, child_hand_seg = FALSE)) %>% ## replace NAs from segmentation with false for hands since all we got annotations for all frames with hands 
  group_by(child_id, age_day_bin) %>%
  summarize(num_detect = length(unique(frame_id)), prop_faces = sum(face_present_ketan) / num_detect, prop_hands = sum(hand_present_ketan) / num_detect, prop_adult_hands = sum(adult_hand_seg) / num_detect, prop_child_hands = sum(child_hand_seg) / num_detect) %>%
  filter(num_detect > 50) %>% ## need at least x samples per point
  mutate(faces_vs_hands = prop_faces - prop_hands, faces_vs_adult_hands = prop_faces - prop_adult_hands, faces_vs_child_hands = prop_faces - prop_child_hands)  %>%
  filter(!is.na(child_id))

```

### Faces, Adult Hands, Child Hands x Age/Child
```{r}
a_hands = ggplot(face_vs_hands_goldset, aes(x=age_day_bin, y=prop_adult_hands, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth(span=10) +
  ylab("Prop adult hands") +
  xlab("Age (weeks)") +
  ylim(0,1) +
  theme(legend.position = "none")

c_hands = ggplot(face_vs_hands_goldset, aes(x=age_day_bin, y=prop_child_hands, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth(span=10) +
  ylab("Prop child hands") +
  xlab("Age (weeks)") + 
  ylim(0,1) +
  theme(legend.position = "none")

faces = ggplot(face_vs_hands_goldset, aes(x=age_day_bin, y=prop_faces, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth(span=10) +
  ylab("Prop faces") +
  xlab("Age (weeks)")
  ylim(0,1)

ggarrange(a_hands, c_hands, faces, nrow=1)

```

### Faces vs Hands (all), Faces vs Adult Hands
```{r}
faces_vs_adult_hands = ggplot(face_vs_hands_goldset, aes(x=age_day_bin, y=faces_vs_adult_hands, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth(span=5) +
  ylab(" faces - adult hands") +
  ylim(-.5,.5)

faces_vs_all_hands = ggplot(face_vs_hands_goldset, aes(x=age_day_bin, y=prop_faces - prop_hands, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth(span=5) +
  ylab(" faces - all hands") +
  ylim(-.5,.5)

ggarrange(faces_vs_adult_hands, faces_vs_all_hands, nrow=1)
```

### Look at any variance by location from single-location videos
```{r}
faces_by_location <- d %>%
  filter(!is.na(Location)) %>%
  filter(count_locations==1) %>%
  group_by(Location, child_id) %>%
  multi_boot_standard(col="face_present_ketan") %>%
  ungroup %>%
  mutate(Location = fct_reorder(Location, mean))

hands_by_location <- d %>%
  filter(!is.na(Location)) %>%
  filter(count_locations==1) %>%
  group_by(Location, child_id) %>%
  multi_boot_standard(col="hand_present_ketan") %>%
  ungroup %>%
  mutate(Location = fct_reorder(Location, mean))

(plot_faces_loc = ggplot(faces_by_location, aes(x = Location, y = mean, col=child_id)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  ylim(0,.8) +
  coord_flip() + 
  ylab('Proportion Faces'))

(plot_hand_loc = ggplot(hands_by_location, aes(x = Location, y = mean, col=child_id)) + 
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper)) + 
  coord_flip() + 
  ylim(0,.8) +
  ylab('Proportion Hands'))

# ggarrange(plot_faces_loc, plot_hand_loc)

```

## Analyze basics from bounding boxes on goldset
### Preprocess bb data
```{r}
bb <- bb %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) %>%
  mutate(x_pos = left*width_px, y_pos = top*height_px) %>%
  mutate(center_x = x_pos + width_x/2, center_y = y_pos - height_y/2) %>%
  filter(!is.na(child_id))

bb_forheatmap <- bb %>%
  filter(height>0)
```

### Number of people, face/hand area, position descriptives
```{r}
##
num_people_goldset <- bb %>%
  group_by(child_id, age_day_bin) %>%
  summarize(num_people = mean(num_faces), num_detect = length(num_faces))

face_area_goldset <- bb %>%
  filter(mean_conf > 0) %>%
  filter(face_openpose==1) %>%
  group_by(child_id, age_day_bin) %>%
  summarize(face_area = mean(area), num_detect = length(area))

hand_area_goldset <- bb %>%
  filter(mean_conf > 0) %>%
  filter(hand_openpose==1) %>%
  group_by(child_id, age_day_bin) %>%
  summarize(hand_area = mean(area), num_detect = length(area))

pos_goldset <- bb %>%
  filter(mean_conf > 0) %>%
  group_by(child_id, age_day_bin, label) %>%
  summarize(x_center = mean(center_x), num_detect = length(center_x))

top_pos_goldset <- bb %>%
   filter(mean_conf > 0) %>%
  group_by(child_id, age_day_bin, label) %>%
  summarize(y_center = mean(center_y), num_detect = length(center_y))

## number of people
num_people_goldset$age_day_bin  = as.numeric(num_people_goldset$age_day_bin)
ggplot(num_people_goldset, aes(x=age_day_bin, y=num_people, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth() +
  ylab("# people")

## face size
face_area_goldset$age_day_bin  = as.numeric(face_area_goldset$age_day_bin)
ggplot(face_area_goldset, aes(x=age_day_bin, y=face_area, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth() +
  ylab("face size")

## hand size
hand_area_goldset$age_day_bin  = as.numeric(hand_area_goldset$age_day_bin)
ggplot(hand_area_goldset, aes(x=age_day_bin, y=hand_area, col=child_id, size=num_detect)) +
  geom_point(alpha=.5) +
  geom_smooth() +
  ylab("hand size")

## y position
top_pos_goldset$age_day_bin  = as.numeric(top_pos_goldset$age_day_bin)
ggplot(top_pos_goldset, aes(x=age_day_bin, y=y_center, col=label)) +
  geom_point(alpha=.5) +
  geom_smooth() +
  ylab("detection_position_y")

```

### Code ot render out bounding boxes from OP on sample frames
```{r}
# g_seg_info <- g_seg %>%
#   distinct(full_image_path, vid_name_short, frame_ind)
#   
# bb_check <- bb %>%
#   mutate(frame_ind = frame, vid_name_short = vid_name) %>%
#   left_join(g_seg_info, by=c("vid_name_short","frame_ind")) %>%
#   filter(!is.na(full_image_path)) %>%
#   filter(mean_conf > 0)

# # dir.create(paste0('bbcheck/'))
# unique_images <- bb_check$full_image_path[101:200]
# 
# for (image in unique_images){
#   this_image_rgb <- image_read(image)
#   img <- image_draw(this_image_rgb)
#   a <- bb_check %>%
#     filter(full_image_path %in% image) 
#   
#   # rect(a$x_pos, a$y_pos-a$height_y, a$x_pos+a$width_x, a$y_pos)
#   text(a$center_x, a$center_y, a$center_y)
#   
#   annotated_file = paste0('bbcheck/',unique(a$vid_name_short), unique(a$frame_ind), 'OP-center.jpg')
#   image_write(img, annotated_file)
#   dev.off()
# }
```

### Heatmaps of face detection positions across age groups
```{r}
face_heatmap_young <- bb_forheatmap %>%
  filter(label=='face') %>%
  filter(age_day_bin<12)

face_heatmap_middle<- bb_forheatmap %>%
  filter(label=='face') %>%
  filter(age_day_bin>12 & age_day_bin<18.1)

face_heatmap_oldest <- bb_forheatmap %>%
  filter(label=='face') %>%
  filter(age_day_bin>18.1)

(p1_youngest <- ggplot(face_heatmap_young, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ggtitle('Faces: 6-12m') +
  ylim(0,height_px) +
  scale_y_reverse())

(p2_middle <- ggplot(face_heatmap_middle, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ggtitle('Faces: 12-18m') +
  ylim(0,height_px) +
  scale_y_reverse())

(p3_oldest <- ggplot(face_heatmap_oldest, aes(x=x_pos, y=y_pos)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ggtitle('Faces: 18m +') +
  ylim(0,height_px) +
  scale_y_reverse())

# ggarrange(p1_youngest, p2_middle,p3_oldest, nrow=1)
```

### Heatmaps of hand detection positions across age groups
```{r}
hand_heatmap_young <- bb_forheatmap %>%
  filter(label=='hand') %>%
  filter(age_day_bin<12)

hand_heatmap_middle<- bb_forheatmap %>%
  filter(label=='hand') %>%
  filter(age_day_bin>12 & age_day_bin<18.1)

hand_heatmap_oldest <- bb_forheatmap %>%
  filter(label=='hand') %>%
  filter(age_day_bin>18.1)

(p1_youngest <- ggplot(hand_heatmap_young, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ggtitle('Hands: 6-12m') +
  ylim(0,height_px) +
  scale_y_reverse())

(p2_middle <- ggplot(hand_heatmap_middle, aes(x=center_x, y=center_y)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ggtitle('Hands: 12-18m') +
  ylim(0,height_px) +
  scale_y_reverse())

(p3_oldest <- ggplot(hand_heatmap_oldest, aes(x=x_pos, y=y_pos)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ggtitle('Hands: 18m +') +
  ylim(0,height_px) +
  scale_y_reverse())

# ggarrange(p1_youngest, p2_middle,p3_oldest, nrow=1)

```

### CHild vs adult hand heatmaps from mturk sample
```{r}

child_hands <- g_seg %>%
  filter(label=="Child hand")

adult_hands <- g_seg %>%
  filter(label=="Adult hand")
##
ggplot(child_hands, aes(x=left, y=top)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ggtitle('Child hands (mturk)') +
  ylim(0,height_px) +
  scale_y_reverse()
##
ggplot(adult_hands, aes(x=left, y=top)) +
  geom_point() + 
  stat_density_2d(aes(fill = ..level..), geom="polygon") +
  coord_fixed(ratio=1) +
  ylim(0,height_px) +
  scale_y_reverse() +
  ggtitle('Adult hands (mturk)')

```