### 2020-01-15 analysis/preprocess_datasets.Rmd

---
title: "saycam full dataset -- preprocessing"
author: "Bria Long"
date: "1/9/2020"
output: html_document
---

# Basic data loading and setup
## Packages
```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
theme_set(theme_few())
```

## Read in data

### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
```{r}
right_side_up_file = "../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

# get INCORRECT videos 
flipped_videos <- right_side_up %>%
  filter(right_side_up==1) %>%
  rename(vid_name = video)
```

### Load SAYCAM metadata from csv file; get metadata from detections .csv
```{r}
metadata_file = "../data/SAYcam-2019-12-0922_54_28.csv"

meta=read.csv(metadata_file) %>%
  rowwise() %>%
  mutate(vid_name = str_split(File.Name,"[.]")[[1]][1]) %>%
  mutate(count_locations = length(str_split(Location,',')[[1]]))

# collapse duplicate labels
meta$Location[meta$Location=='Bedroom Alice/Sam'] = 'Bedroom'

```

### Load in detections & bounding boxes from preprocessed Rdata file; 
```{r eval=FALSE}
 # Could do list of detections  from csv (takes forever, skipping)
# data_file = "../data/all_frames.csv"
# detections=read.csv(data_file)

## rdata is faster
load('../data/openpose_detections.RData')

```

### Preprocess data (doesn't execute in markdown; loads processed data below)
```{r eval=FALSE}
# get shortened video names so they match across all files
all_video_names = as_tibble(unique(detections$vid_name))
all_video_names <- all_video_names %>%
  rowwise() %>%
  mutate(vid_name = str_split(value,"[.]")[[1]][1])

# reset levels of vid_name in big detections array for detections & bb (very slow rowwise)
levels(detections$vid_name) <- all_video_names$vid_name

# join with metadata and exclude bad videos
d_all <- detections %>%
  left_join(meta, by='vid_name') %>%
  filter(!vid_name %in% flipped_videos$vid_name)

## All 6 of the video that were not in this file (but were in the detections file) happened to be 6 very bad quality videos; excluding from detections
exclude_bad_quality <- d_all %>%
  filter(!vid_name %in% right_side_up$video)

d_all <- d_all %>%
  filter(!vid_name %in% exclude_bad_quality$vid_name)

# check that we actually excluded the videos
num_orig_videos = length(unique(detections$vid_name))
num_videos_retained = length(unique(d_all$vid_name))
assert_that(num_videos_retained < num_orig_videos)

# get rid of big data structure
remove(detections)
```

### Make detections into logical vectors and compute "person" as either face/hand
```{r}
d_all$face_openpose = as.logical(d_all$face_openpose)
d_all$hand_openpose = as.logical(d_all$hand_openpose)
d_all$person_openpose = (as.logical(d_all$face_openpose) | as.logical(d_all$hand_openpose))
d_all$both_openpose = (as.logical(d_all$face_openpose) & as.logical(d_all$hand_openpose))
```

### Look at any suspicious videos with very high face detections (candidate missed allocentric videos)
```{r}
vid_check <- d_all %>%
  filter(child_id == "S") %>%
  group_by(vid_name,vid_name, Databrary.Link) %>%
  summarize(prop_faces =mean(face_openpose))

high_faces <- vid_check %>%
  filter(prop_faces>.8)
```

### Make bins across which to analyze data
```{r}
bin_size = 7
min_age = min(d_all$age_days)
max_age = max(d_all$age_days)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)
```

### And save out into dataframe
```{r}
d_all <- d_all %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1)))
d_all$age_day_bin = as.numeric(as.character(d_all$age_day_bin))
```

### Save this data so this doesn't have to be redone again
```{r}
save(d_all, file = '../data/openpose_detections_filtered.RData')
```

## Load bounding boxes to get cropped detections
```{r}
## import bounding boxes
load('../data/all_bounding_boxes.RData')

## frame
height_px = 480; width_px = 640

bb_all <- df %>%
  filter(right_side_up==0) %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin)))

## get rid of extra dataframe
remove(df)
```

## Make cropped detections in top 60% of frame
```{R}
detections_cropped <- bb_all %>%
  filter(height <.6) %>%  ## cut off bottom part of frame
  group_by(vid_name, frame, child_id, age_day_bin) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0)) %>% # if any faces or hands
  distinct(vid_name, frame, age_day_bin,face_detected, hand_detected)

## clean up so that we can join with original dataframe
detections_cropped <- detections_cropped %>%
  ungroup() %>%
  rename(hand_detected_cropped = hand_detected, face_detected_cropped = face_detected) %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id))

## now join with big dataframe
d_cropped <- d_all %>%
  as_tibble() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id),face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
  select(vid_name, frame, child_id, age_day_bin, face_openpose, hand_openpose) %>%
  left_join(detections_cropped, by=c("vid_name","frame","age_day_bin","child_id"))

d_cropped <- d_cropped %>%
  replace_na(list(hand_detected_cropped = FALSE, face_detected_cropped = FALSE))  ## replace NAs from cropped anlaysis

# and save it out
save(d_cropped, file = '../data/openpose_detections_filtered_cropped.RData')
```

# Render out images based on annotations if desired
## True positives on faces?
```{r eval=FALSE}
faces_only_op_cor <- m %>%
  filter(faces == TRUE) %>%
  filter(face_openpose == TRUE)

# how about images where open pose got it? these should be most obvious
dir.create(paste0('faces_tp/'))
for (i in seq(1,length(faces_only_op_cor$full_image_path),1)){
  image_read(as.character(faces_only_op_cor$full_image_path[i])) %>%
  image_append(stack = FALSE) %>%
  image_write(file.path(paste0("faces_tp/", faces_only_op_cor$short_image_path[i])))
}

```

## False positives on facs?
```{r eval=FALSE}
to_write <- m %>%
  filter(faces == FALSE) %>%
  filter(face_openpose == TRUE)

# how about images where open pose got it? these should be most obvious
dir.create(paste0('faces_fp/'))
for (i in seq(1,length(to_write$full_image_path),1)){
  image_read(as.character(to_write$full_image_path[i])) %>%
  image_append(stack = FALSE) %>%
  image_write(file.path(paste0("faces_fp/", to_write$short_image_path[i])))
}

```

## False negatives on faces?
```{r eval=FALSE}
to_write <- m %>%
  filter(faces == TRUE) %>%
  filter(face_openpose == FALSE)

# how about images where open pose got it? these should be most obvious
dir.create(paste0('faces_fn/'))
for (i in seq(1,length(to_write$full_image_path),1)){
  image_read(as.character(to_write$full_image_path[i])) %>%
  image_append(stack = FALSE) %>%
  image_write(file.path(paste0("faces_fn/", to_write$short_image_path[i])))
}

```