### 2020-01-24 analysis/preprocess_datasets.Rmd

---
title: "saycam full dataset -- preprocessing"
author: "Bria Long"
date: "1/9/2020"
output: html_document
---

# Basic data loading and setup
## Packages
```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
theme_set(theme_few())
```

## Read in data

### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
```{r}
right_side_up_file = "../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

# get INCORRECT videos 
flipped_videos <- right_side_up %>%
  filter(right_side_up==1) %>%
  rename(vid_name = video)
```

### Load SAYCAM metadata from csv file; get metadata from detections .csv
```{r}
metadata_file = "../data/SAYcam-2019-12-0922_54_28.csv"

meta=read.csv(metadata_file) %>%
  rowwise() %>%
  mutate(vid_name = str_split(File.Name,"[.]")[[1]][1]) %>%
  mutate(count_locations = length(str_split(Location,',')[[1]]))

# collapse duplicate labels
meta$Location[meta$Location=='Bedroom Alice/Sam'] = 'Bedroom'

save(meta, file = '../data/saycam_metadata.RData')

```

### Load in detections & bounding boxes from preprocessed Rdata file; 
```{r eval=FALSE}
 # Could do list of detections  from csv (takes forever, skipping)
# data_file = "../data/all_frames.csv"
# detections=read.csv(data_file)

## rdata is faster
load('../data/openpose_detections.RData')

```

### Preprocess data (doesn't execute in markdown; loads processed data below)
```{r eval=FALSE}
# get shortened video names so they match across all files
all_video_names = as_tibble(unique(detections$vid_name))
all_video_names <- all_video_names %>%
  rowwise() %>%
  mutate(vid_name = str_split(value,"[.]")[[1]][1])

# reset levels of vid_name in big detections array for detections & bb (very slow rowwise)
levels(detections$vid_name) <- all_video_names$vid_name

# join with metadata and exclude bad videos
d_SA <- detections %>%
  left_join(meta, by='vid_name') %>%
  filter(!vid_name %in% flipped_videos$vid_name)

## All 6 of the video that were not in this file (but were in the detections file) happened to be 6 very bad quality videos; excluding from detections
exclude_bad_quality <- d_SA %>%
  filter(!vid_name %in% right_side_up$video)

d_SA <- d_SA %>%
  filter(!vid_name %in% exclude_bad_quality$vid_name)

# check that we actually excluded the videos
num_orig_videos = length(unique(detections$vid_name))
num_videos_retained = length(unique(d_SA$vid_name))
assert_that(num_videos_retained < num_orig_videos)

# get rid of big data structure
remove(detections)
```

## Load, process, merge Y's data
### Now load in bounding box data from Y, which has detection rates built in
```{r}
load('../data/Y_bounding_boxes.RData') #df again
y_data <- df %>%
  rowwise() %>%
  mutate(age = str_split_fixed(vid_name, '_',4)[,3])

remove(df)

### wrangle age data for Y
d_y <- y_data %>%
  mutate(age_months = as.numeric(substr(age, 0,2)), days_to_add = as.numeric(substr(age, 3,4))) %>%
  mutate(age_days = age_months*30 + days_to_add)

save(y_data, file = '../data/Y_bounding_boxes_cleaned.RData')
```

### Do basic preprocessing on Y detection's
```{r}

### count hands/faces from bounding box outputs for Y -- for d_all
y_detections <- d_y %>%
  group_by(vid_name, frame,age_days) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0))

y_detections <- y_detections %>%
  # rename(hand_openpose = hand_detected, face_openpose = face_detected) %>%
  mutate(child_id = "Y") %>%
  left_join(meta)

save(y_detections, file='../data/y_detections_preprocessed.RData')
```

### Merge detections from Y with other kids
```{r}
d_all <- d_SA %>%
  mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
  full_join(y_detections)
```

### Save this data so this doesn't have to be redone again
```{r}
save(d_all, file = '../data/openpose_detections_filtered_threekids.RData')
```
## Save out vid info for use with joining stuff later
```{r}
# vid_info_all <- d_all %>%
#   select(vid_name, frame, age_days, child_id)
# 
# save(vid_info_all, '../data/vid_info_all.RData')
```
### Look at any suspicious Sam videos with very high face detections (candidate missed allocentric videos)
```{r eval=FALSE}
# vid_check <- d_all %>%
#   filter(child_id == "S") %>%
#   group_by(vid_name,vid_name, Databrary.Link) %>%
#   summarize(prop_faces =mean(face_openpose))
# 
# high_faces <- vid_check %>%
#   filter(prop_faces>.8)
```

# Cropped detections
## Load bounding boxes from S/A to get cropped detections
```{r}
## import bounding boxes
load('../data/all_bounding_boxes.RData')

## frame
height_px = 480; width_px = 640

bb_all <- df %>%
  filter(right_side_up==0)

## get rid of extra dataframe
remove(df)
```

## Make cropped detections in top 60% of frame for Y
```{r}
load(file = '../data/Y_bounding_boxes_cleaned.RData')
d_y <- y_data %>%
  mutate(age_months = as.numeric(substr(age, 0,2)), days_to_add = as.numeric(substr(age, 3,4))) %>%
  mutate(age_days = age_months*30 + days_to_add)

detections_cropped_Y <- d_y %>%
  filter(person==0) %>% #only get frames where detections
  filter(height <(.6*480)) %>%  ## cut off bottom part of frame
  group_by(vid_name, frame, child_id, age_days) %>%
  summarize(hand_detected_cropped = (sum(label=="hand")>0), face_detected_cropped = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))

detections_cropped_Y <- detections_cropped_Y %>%
 ungroup() %>%
 mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id))

```

## Make cropped detections in top 60% of frame for S,A
```{R}
detections_cropped_SA <- bb_all %>%
  filter(height <.6) %>%  ## cut off bottom part of frame
  group_by(vid_name, frame, child_id, age_days) %>%
  summarize(hand_detected_cropped = (sum(label=="hand")>0), face_detected_cropped = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))

all_cropped_detections <- detections_cropped_SA %>%
  ungroup() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  full_join(detections_cropped_Y)

```

## need to get "FALSE" on frames where we didn't have info to be able to compute proportions; uses d_all from earlier and fills it all back in
```{r}
# load('../data/openpose_detections_filtered_threekids.RData') # d_all

d_cropped <- d_all %>% ## needs to be loaded
  as_tibble() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  select(vid_name, frame, child_id, age_days, face_openpose) %>%
  left_join(all_cropped_detections, by=c("vid_name","frame","age_days","child_id")) %>%
  replace_na(list(hand_detected_cropped = FALSE, face_detected_cropped = FALSE ))
```

### now save
```{r}
save(d_cropped, file = '../data/openpose_detections_filtered_cropped_allthreekids.RData')

```

# Make bb detections only
```{r}
## import bounding boxes
load('../data/all_bounding_boxes.RData')
load('../data/Y_bounding_boxes_cleaned.RData')

right_side_up_file = "../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

# get INCORRECT videos 
flipped_videos <- right_side_up %>%
  filter(right_side_up==1) %>%
  rename(vid_name = video)

```

```{r}
## frame
height_px = 480; width_px = 640
bb_all <- y_data %>%
  filter(person==0) %>% # only rows with detections for now
  filter(!is.na(child_id)) %>%
  mutate(left = left*width_px, top = top*height_px, height = height*height_px, width = width*width_px) %>% #unnormalize data...
  full_join(df) %>% # join with S/A
  filter(right_side_up==0) %>% # get rid of allocentric
  mutate(center_x = left + width/2, center_y = top - height/2, area = width*height)

save(bb_all, file = '../data/all_bounding_boxes_three_kids_preprocessed.RData')

```

```