### 2020-01-28 analysis/preprocess_datasets.Rmd
fix merging datasets bug in preprocessing pipelined

---
title: "saycam full dataset -- preprocessing"
author: "Bria Long"
date: "1/9/2020"
output: html_document
---

# Basic data loading and setup
## Packages
```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
theme_set(theme_few())
```

## Read in data

### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
```{r}
right_side_up_file = "../data/video_right-side-up.csv"
right_side_up =read.csv(right_side_up_file)

# get INCORRECT videos 
flipped_videos <- right_side_up %>%
  filter(right_side_up==1) %>%
  rename(vid_name = video)
```

### Load SAYCAM metadata from csv file; get metadata from detections .csv
```{r}
# metadata_file = "../data/SAYcam-2019-12-0922_54_28.csv"
# meta=read.csv(metadata_file) %>%
#   rowwise() %>%
#   mutate(vid_name = str_split(File.Name,"[.]")[[1]][1]) %>%
#   mutate(count_locations = length(str_split(Location,',')[[1]]))
# 
# # collapse duplicate labels
# meta$Location[meta$Location=='Bedroom Alice/Sam'] = 'Bedroom'
# save(meta, file = '../data/saycam_metadata.RData')
load(file = '../data/saycam_metadata.RData')
```

### Load in detections & bounding boxes from preprocessed Rdata file; 
```{r}
 # Could do list of detections  from csv (takes forever, skipping)
# data_file = "../data/all_frames.csv"
# detections=read.csv(data_file)

## rdata is faster
load('../data/openpose_detections.RData')
```

### Preprocess data (doesn't execute in markdown; loads processed data below)
```{r}
# get shortened video names so they match across all files
all_video_names = as_tibble(unique(detections$vid_name))
all_video_names <- all_video_names %>%
  rowwise() %>%
  mutate(vid_name = str_split(value,"[.]")[[1]][1])

# reset levels of vid_name in big detections array for detections & bb (very slow rowwise)
levels(detections$vid_name) <- all_video_names$vid_name

# join with metadata and exclude bad videos
d_SA <- detections %>%
  left_join(meta, by='vid_name') %>%
  filter(!vid_name %in% flipped_videos$vid_name)

## All 6 of the video that were not in this file (but were in the detections file) happened to be 6 very bad quality videos; excluding from detections
exclude_bad_quality <- d_SA %>%
  filter(!vid_name %in% right_side_up$video)

d_SA <- d_SA %>%
  filter(!vid_name %in% exclude_bad_quality$vid_name)

# check that we actually excluded the videos
num_orig_videos = length(unique(detections$vid_name))
num_videos_retained = length(unique(d_SA$vid_name))
assert_that(num_videos_retained < num_orig_videos)

# get rid of big data structure
remove(detections)
```

## Load, process, merge Y's data
### Now load in bounding box data from Y, which has detection rates built in
```{r}
# load('../data/Y_bounding_boxes.RData') #df again
# y_data <- df %>%
#   rowwise() %>%
#   mutate(age = str_split_fixed(vid_name, '_',4)[,3])
# remove(df)
# 
# ### wrangle age data for Y
d_y <- y_data %>%
  mutate(age_months = as.numeric(substr(age, 0,2)), days_to_add = as.numeric(substr(age, 3,4))) %>%
  mutate(age_days = age_months*30 + days_to_add)
## 
save(d_y, file = '../data/Y_bounding_boxes_cleaned_jan27.RData')
```

### Do basic preprocessing on Y detection's
```{r}
### count hands/faces from bounding box outputs for Y -- for d_all
# y_detections <- d_y %>%
#   group_by(vid_name, frame,age_days) %>%
#   summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0))
# 
# y_detections <- y_detections %>%
#   mutate(child_id = "Y") %>%
#   left_join(meta)
#   
# save(y_detections, file='../data/y_detections_preprocessed.RData')
```

### Merge detections from Y with other kids
```{r}
load(file='../data/y_detections_preprocessed.RData')
y_detections <-  y_detections %>%
  rename(hand_openpose = hand_detected, face_openpose = face_detected) # standardize

d_all <- d_SA %>%
  mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
  full_join(y_detections)

# check
length(unique(d_all$vid_name)) == (length(unique(y_detections$vid_name)) +  length(unique(d_SA$vid_name)))

# clean up
remove(d_SA)
remove(d_y)

# save
save(d_all, file = '../data/openpose_detections_filtered_threekids_jan27.RData')
```

## Save out vid info for use with joining stuff later
```{r}
# vid_info_all <- d_all %>%
#   select(vid_name, frame, age_days, child_id)
# 
# save(vid_info_all, '../data/vid_info_all.RData')
```
### Look at any suspicious Sam videos with very high face detections (candidate missed allocentric videos)
```{r eval=FALSE}
# vid_check <- d_all %>%
#   filter(child_id == "S") %>%
#   group_by(vid_name,vid_name, Databrary.Link) %>%
#   summarize(prop_faces =mean(face_openpose))
# 
# high_faces <- vid_check %>%
#   filter(prop_faces>.8)
```

# Cropped detections
## Load bounding boxes from S/A to get cropped detections
```{r}
## import bounding boxes
load('../data/all_bounding_boxes.RData')
bb_all <- df
remove(df)

## frame
height_px = 480; width_px = 640

# get shortened video names so they match across all files
all_video_names_bb = as_tibble(unique(bb_all$vid_name))
all_video_names_bb <- all_video_names_bb %>%
  rowwise() %>%
  mutate(vid_name = str_split(value,"[.]")[[1]][1])

# reset levels of vid_name in big detections array for detections & bb (very slow rowwise)
bb_all$vid_name <- as.factor(bb_all$vid_name)
levels(bb_all$vid_name) <- all_video_names_bb$vid_name

bb_all <- bb_all %>%
  filter(!vid_name %in% flipped_videos$video)

save(bb_all, file = "../data/SA_bounding_boxes_cleaned.RData")

# length(unique(bb_all_2$vid_name[bb_all_2$child_id=='S']))
# length(unique(bb_all_2$vid_name[bb_all_2$child_id=='A']))
# 
# length(unique(d_all$vid_name[d_all$child_id=='S']))
# length(unique(d_all$vid_name[d_all$child_id=='A']))
```

## Join detections for both and save out for later
```{r}
# load(file = '../data/Y_bounding_boxes_cleaned.RData') #  d_y from earlier if not loaded

## frame
# height_px = 480; width_px = 640
# bb_all_three <- d_y %>%
#   filter(person==0) %>% # only rows with detections for now so comparable to bb_all
#   mutate(left = left*width_px, top = top*height_px, height = height*height_px, width = width*width_px) %>% #unnormalize data so comparable
#   full_join(bb_all) %>% # join with SA bounding boxes
#   mutate(center_x = left + width/2, center_y = top - height/2, area = width*height) 
# 
# save(bb_all_three, file = '../data/all_bounding_boxes_three_kids_preprocessed.RData')
# remove(bb_all_three)
```

```{r}
load(file='../data/openpose_detections_filtered_threekids_jan27.RData')
load(file='../data/SA_bounding_boxes_cleaned.RData')
load(file='../Y_bounding_boxes_cleaned_jan27.RData')
```

### Make cropped detections for Y/SA, join, make cropped large data
```{r}
detections_cropped_Y <- d_y %>%
  filter(person==0) %>% #only get frames where detections so comparable
  filter((height<(.6*480))) %>%  ## 
  group_by(vid_name, frame, child_id, age_days) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))

detections_cropped_Y <- detections_cropped_Y %>%
 ungroup() %>%
 mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id))

## Make cropped detections in top 60% of frame for S,A
detections_cropped_SA <- bb_all %>%
  filter(height<(.6)) %>%  ## 
  group_by(vid_name, frame, child_id, age_days) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))  
# join all together
all_cropped_detections <- detections_cropped_SA %>%
  ungroup() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  full_join(detections_cropped_Y)

## clean up workspace or else we run into memory limits
# cropped detections
remove(detections_cropped_Y)
remove(detections_cropped_SA)

# bb data
remove(y_data)
remove(bb_all)

## need to get "FALSE" on frames where we didn't have info to be able to compute proportions; uses d_all from earlier and fills it all back in

all_cropped_detections <- all_cropped_detections %>%
  select(-age_days, -child_id)

d_cropped <- d_all %>% ## needs to be loaded!
  as_tibble() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  select(vid_name, frame, child_id, age_days, face_openpose, hand_openpose) %>%
  left_join(all_cropped_detections, by=c("vid_name","frame"))

save(d_cropped, file = '../data/openpose_detections_filtered_cropped_allthreekids_jan27.RData')

```

```{r}
what <- d_cropped %>%
  filter(face_openpose == TRUE) %>%
  filter(fa)

vid_summary <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  group_by(vid_name) %>%
  summarize(prop_faces = mean(face_openpose), prop_faces_cropped = mean(face_detected), prop_hands_cropped = mean(hand_detected), diff = prop_faces - prop_faces_cropped) %>%
  filter(diff>.2)

weird_vids = unique(vid_summary$vid_name[vid_summary$prop_hands_cropped==0])

weird_d_all <- d_all %>%
  filter(vid_name %in% weird_vids)

bb_all_check <- bb_all %>%
  filter(vid_name %in% weird_vids)
```