### 2020-01-30 analysis/analysis_bb_fulldataset_2.Rmd
code to start bb analyses in full data, part2

---
title: "analyze_bb_fullset"
author: "Bria Long"
date: "1/23/2020"
output: html_document
---

# Basic data loading and setup
## Packages
```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(egg)
theme_set(theme_few())
```

### Load preprocessed bb's for all kids
```{r}
load(file = '../data/all_bounding_boxes_three_kids_preprocessed.RData')

```

### Make bins across which to analyze data
```{r}
bin_size = 7
min_age = min(bb_all$age_days)
max_age = max(bb_all$age_days)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)

bb_all <- bb_all %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(d$age_day_bin)))
# 
# bin_size = 6
# min_age = 6
# max_age = 30
# bin_starts = seq(min_age, max_age+1,bin_size)
# bins = c(bin_starts)
```

```{r}
hands_by_size_by_frame <- bb_all %>%
  filter(label=="hand") %>%
  group_by(vid_name, frame, age_days, child_id) %>%
  summarize(num_hands = sum(label=="hand"), avg_area = mean(area), avg_x = mean(center_x), avg_y = mean(center_y))

faces_by_size_by_frame <- bb_all %>%
  filter(label=="face") %>%
  group_by(vid_name, frame, age_days, child_id) %>%
  summarize(num_faces = sum(label=="face"), avg_area = mean(area), avg_x = mean(center_x), avg_y = mean(center_y))

poses_by_size_by_frame <- bb_all %>%
  filter(label=="pose") %>%
  group_by(vid_name, frame, age_days, child_id) %>%
  summarize(num_poses = sum(label=="pose"), avg_area = mean(area), avg_x = mean(center_x), avg_y = mean(center_y))
```

```{r}

to_plot_poses <- poses_by_size_by_frame %>%
  mutate(age_months = age_days/30) %>%
  mutate(months_binned = cut(age_months, bins, labels=c("6m-12m","12m-18m","18m-24m","24m-32m")))

save(hands_by_size_by_frame, file="../data/hands_by_size_by_frame.RData")
save(faces_by_size_by_frame, file="../data/faces_by_size_by_frame.RData")
save(poses_by_size_by_frame, file="../data/poses_by_size_by_frame.RData")

```

```{r}
bin_size = 7
min_age = min(bb_all$age_days)
max_age = max(bb_all$age_days)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)

bb_all <- bb_all %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(d$age_day_bin)))

poses_by_size_by_frame <- poses_by_size_by_frame %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(d$age_day_bin)))

```

```{r}
ggplot(to_plot_poses, aes(x=num_poses, y=log(avg_area), col=child_id)) +
  geom_jitter(alpha=.2) +
  ggtitle('Pose size by #') +
  geom_grid(~months_binned)

ggplot(to_plot_hands, aes(x=num_hands, y=log(avg_area), col=child_id)) +
  geom_jitter(alpha=.2) +
  ggtitle('Hand size by #') +
  geom_grid(~months_binned)

ggplot(to_plot_faces, aes(x=num_faces, y=log(avg_area), col=child_id)) +
  geom_jitter(alpha=.2) +
  ggtitle('Face size by #') +
  geom_grid(~months_binned)

```

```{r}
summary_by_age <- bb_all %>%
  group_by(age_day_bin, child_id) %>%
  summarize(num_detect = length(face_openpose), prop_faces = sum(face_openpose) / num_detect,  prop_hands = sum(hand_openpose) / num_detect, face_or_hand = sum(person_openpose)/ num_detect, face_and_hand = sum(both_openpose)/ num_detect)

```