### 2021-04-16 analysis/_old_analysis/bb_check.Rmd
reorg analysis scripts -- port outdated/draft scripts to "old" folder

---
title: "BB_CHeck"
author: "Bria Long"
date: "5/2/2020"
output: html_document
---

# Load temp files if we are skipping steps above run if we are skipping steps above
```{r eval=FALSE}
## For cropped detections, need..
# 1. full set of detections (d_all from above)
# load(file='../data/preprocessed_data/openpose_detections_filtered_2020-01-29.RData')

# 2. bounding boxes from S and A (bb_SA)
load(file='../data/temp_data/SA_bounding_boxes_cleaned_2020-01-29.RData')

# 3. bounding boxes from Y (d_y) -- has different format than bb_SA, so diff name..
# load(file='../data/temp_data/Y_bounding_boxes_cleaned_2020-01-29.RData')
```

```{r}
load(file='../data/preprocessed_data/openpose_detections_filtered_2020-01-29.RData')
d_SA <- d_all %>%
  filter(child_id %in% c('S','A'))
```

```{r}
detections_cropped_SA <- bb_SA %>%
  filter(height<(.6)) %>%  ## 
  group_by(vid_name, frame, child_id) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"), mean_height = mean(height), mean_top = mean(top))

detections_SA <- bb_SA %>%
  group_by(vid_name, frame, child_id) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"), mean_height = mean(height), mean_top = mean(top))

```

```{r}
hands_bb <- detections_SA %>%
  filter(hand_detected==TRUE)

hands_dets <- d_SA %>%
  filter(hand_openpose==TRUE)

hands_compare <- hands_dets %>%
  left_join(hands_bb)

lacking_bb <- hands_compare %>%
  filter(is.na(hand_detected))

compare_one_vid_BB <- detections_SA %>%
  filter(vid_name=="A_20130531_0818_01")

compare_one_vid_dets <- d_SA %>%
  filter(vid_name=="A_20130531_0818_01")

joined_SA <- d_SA %>%
  left_join(detections_SA) %>%

no_bb <-  joined_SA %>%
  filter(hand_openpose==TRUE & hand_detected==FALSE)

```

## Look at height (i.e., vertical size) of hands
```{r}
bb_hands_y <- d_y %>%
   filter(label=='hand')

bb_hands <- bb_SA %>%
  filter(label=='hand')

bb_faces<- bb_SA %>%
  filter(label=='face')

```

```{r}
what <- bb_hands %>%
  filter(height>.6)
```