### 2021-04-16 analysis/_old_analysis/vid_check.Rmd

---
title: "Untitled"
author: "Bria Long"
date: "7/18/2020"
output: html_document
---

```{r}
library(tidyverse)
library(here)
load(paste0(here::here(), '/data/raw_data/SA_bounding_boxes.RData'))
```

## calculate problems
```{r}
reload_data=TRUE

if (reload_data){
vid_index = 0

for (vid_file in op_files[1700:1710]) {
  vid_index = vid_index + 1

  vid_file_path_new= paste0(here::here(), '/data/op_outputs_per_vid_full_set/', vid_file)
  this_vid_data_new <- readRDS(vid_file_path_new)

   vid_file_path_old = paste0(here::here(), '/data/_old_op_outputs_per_vid/', vid_file)
   this_vid_data_old <- readRDS(vid_file_path_old)

   # check vid name to see if excluded vid
   this_vid_name_temp <- this_vid_data_new$vid_name[1]
   this_vid_name <- str_split(this_vid_name_temp,"[.]")[[1]][1]

   # only run if not excluded
   if (!this_vid_name %in% to_exclude$vid_name){

    assert_that(this_vid_data_new$vid_name[1] == this_vid_data_old$vid_name[1])
    # summarize hands/faces in each video
   summary_vid_data_new <- this_vid_data_new %>%
     group_by(frame, child_id, vid_name) %>%
     summarize(hand_detected = sum(!is.na(mean_confidence[label=='hand']))>0,face_detected = sum(!is.na(mean_confidence[label=='face']))>0)  %>%
     group_by(child_id, vid_name) %>%
     summarize(num_hands = sum(hand_detected), num_faces = sum(face_detected), num_frames = length(hand_detected))

      summary_vid_data_old <- this_vid_data_old %>%
     group_by(frame, child_id, vid_name) %>%
     summarize(hand_detected = sum(!is.na(mean_confidence[label=='hand']))>0,face_detected = sum(!is.na(mean_confidence[label=='face']))>0)  %>%
     group_by(child_id, vid_name) %>%
     summarize(num_hands = sum(hand_detected), num_faces = sum(face_detected), num_frames = length(hand_detected))

      video_diff <- summary_vid_data_new %>%
        mutate(diff_hands = num_hands - summary_vid_data_old$num_hands,
               diff_faces = num_faces - summary_vid_data_old$num_faces) %>%
        mutate(vid_name = vid_file)

  print(video_diff)

  # join it: 1 row = 1 video
   if (vid_index == 1 ){
     all_vid_diffs <- video_diff 
   }
   else if (vid_index > 1) {
   all_vid_diffs <- all_vid_diffs %>%
     full_join(video_diff)
   }
}
}
}
```

```{r}
old_no_na <- this_vid_data_old %>%
  filter(person<3) %>%
  filter(!is.na(mean_confidence))

new_no_na <- this_vid_data_new %>%
  filter(person<3) %>%
  filter(!is.na(mean_confidence))
```

```{r}
files_to_check = op_files[1:100]

all_vid_diffs <- all_vid_diffs %>%
  mutate(vid_name = str_split_fixed(vid_name, '_BBs.Rds',2)[,1])

orig_dectections_check <- d_all %>%
  filter(vid_name %in% all_vid_diffs$vid_name)

orig_dectections_check_by_vid <- orig_dectections_check %>%
  group_by(vid_name) %>%
  summarize(num_faces_orig_det = sum(face_openpose))

all_vid_diffs <- all_vid_diffs %>%
  left_join(orig_dectections_check_by_vid)
```

```{r}
first_vid_faces <- orig_dectections_check %>%
  filter(face_openpose==TRUE) %>%
  filter(vid_name == 'A_20130531_0818_01')

first_vid_faces_new_bb <- this_vid_data_new %>%
  filter(label=='face')

first_vid_faces_old_bb <- this_vid_data_old %>%
  filter(label=='face') %>%
  filter(!is.na(mean_confidence))

first_vid_faces_old_bb_all_NAs <- this_vid_data_old %>%
  filter(label=='face')

```

```{r}

face_frames <- summary_vid_data %>%
  filter(face_detected==TRUE)

frame_level_faces <- this_vid_data %>%
  right_join(face_frames) %>%
  filter(label=='face') %>%
  filter(!is.na(mean_confidence))

```

```{r}

old_faces <- this_vid_data_old %>%
  filter(label=='face') %>%
  filter(!is.na(mean_confidence)) %>%
  filter(person==0)

new_faces <- this_vid_data_new %>%
  filter(label=='face') %>%
  filter(!is.na(mean_confidence)) %>%
  filter(person==0)

```

```{r}
this_vid_name = 'A_20130531_0818_01'

test_vid <- df %>%
  filter(vid_name == this_vid_name)

summary_old_bbs_data <- test_vid %>%
  group_by(frame) %>%
    summarize(hand_detected = sum(!is.na(mean_confidence[label=='hand']))>0,face_detected = sum(!is.na(mean_confidence[label=='face']))>0)  %>%
     summarize(num_hands = sum(hand_detected), num_faces = sum(face_detected))

new_file_path = paste0(here::here(), '/data/op_outputs_per_vid_full_set/', this_vid_name, '_BBs.Rds')
this_vid_data_new <- readRDS(new_file_path)

summary_new_bbs_data <- this_vid_data_new %>%
  group_by(frame) %>%
     summarize(hand_detected = sum(!is.na(mean_confidence[label=='hand']))>0,face_detected = sum(!is.na(mean_confidence[label=='face']))>0)  %>%
     summarize(num_hands = sum(hand_detected), num_faces = sum(face_detected))

summary_new_bbs_data - summary_old_bbs_data
```

```{r}

> s = read_csv('sherlock.csv', col_names=FALSE)

s <- s %>%

```