### 2021-04-16 preprocessing/_old/4_preprocess_datasets.Rmd

---
title: "saycam full dataset -- preprocessing"
author: "Bria Long"
date: "1/9/2020"
output: html_document
---

# Basic data loading and setup
## Packages
```{r setup, include=FALSE}
 knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(magick)
library(stringr)
library(lubridate)
library(egg)
theme_set(theme_few())
## are we saving outputs
saveFiles = TRUE
```

## Read in data

### Load file to filter flipped videos (when sam's mom was wearing camera; not egocentric)
```{r}
allocentric_file = "../data/raw_data/SAYCAM_allocentric_videos.csv"
to_exclude =read.csv(allocentric_file)

# get INCORRECT videos 
to_exclude <- to_exclude %>%
  filter(allocentric==1) %>%
  rename(vid_name = video)
```

### Load SAYCAM metadata from csv file; get metadata from detections .csv
```{r}
metadata_file = "../data/raw_data/Metadata_SAYcam-2019-12-0922_54_28.csv"
meta=read.csv(metadata_file) %>%
  rowwise() %>%
  mutate(vid_name = str_split(File.Name,"[.]")[[1]][1]) %>%
  mutate(count_locations = length(str_split(Location,',')[[1]]))

# collapse duplicate labels
meta$Location[meta$Location=='Bedroom Alice/Sam'] = 'Bedroom'

# only select relevant columnts
meta <- meta %>%
  select(vid_name, count_locations, Location, Databrary.Link)
```

### Load in detections & bounding boxes from preprocessed Rdata file; 
```{r}
 # Could do list of detections  from csv (takes forever, skipping)
# data_file = "../data/raw_data/all_frames.csv"
# detections=read.csv(data_file)

## rdata is faster
load('../data/raw_data/openpose_detections_SA.RData') #SA = for Sam and Alice
# unselect unused columns
detections <- detections %>%
  select(-unix_time, -X, -vid_path, -nose_conf, -wrist_conf)
```

### Preprocess detections 
```{r}
# get shortened video names so they match across all files
all_video_names = as_tibble(unique(detections$vid_name))
all_video_names <- all_video_names %>%
  rowwise() %>%
  mutate(vid_name = str_split(value,"[.]")[[1]][1])

# reset levels of vid_name in big detections array for detections & bb (very slow rowwise)
levels(detections$vid_name) <- all_video_names$vid_name

# join with metadata and exclude bad videos
d_SA <- detections %>%
  left_join(meta, by='vid_name') %>%
  filter(!vid_name %in% to_exclude$vid_name)

## All 6 of the video that were not in this file (but were in the detections file) happened to be 6 very bad quality videos; excluding from detections
exclude_bad_quality <- d_SA %>%
  filter(!vid_name %in% right_side_up$video) %>%
  distinct(vid_name)

d_SA <- d_SA %>%
  filter(!vid_name %in% exclude_bad_quality$vid_name)

# make new list for later
vids_to_exclude <- to_exclude %>%
  select(vid_name) %>%
  full_join(exclude_bad_quality) %>%
  select(vid_name) %>%
  distinct(vid_name)

# check that we actually excluded the videos
num_orig_videos = length(unique(detections$vid_name))
num_videos_retained = length(unique(d_SA$vid_name))
assert_that(num_videos_retained < num_orig_videos)

# get rid of big data structure
remove(detections)
```

## Load, process, merge Y's data
### Now load in bounding box data from Y, which has detection rates built in
```{r}
load('../data/raw_data/Y_bounding_boxes.RData') #df again as import
y_data <- df %>%
  rowwise() %>%
  mutate(age = str_split_fixed(vid_name, '_',4)[,3])
remove(df)

# ### wrangle age data for Y
d_y <- y_data %>%
  mutate(age_months = as.numeric(substr(age, 0,2)), days_to_add = as.numeric(substr(age, 3,4))) %>%
  mutate(age_days = age_months*30.42 + days_to_add)
##
if (saveFiles){
save(d_y, file = paste0('../data/temp_data/Y_bounding_boxes_cleaned_',today(),'.RData'))
}
```

### Do basic preprocessing on Y detection's
```{r}
### count hands/faces from bounding box outputs for Y -- for d_all
y_detections <- d_y %>%
  group_by(vid_name, frame,age_days) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0))

y_detections <- y_detections %>%
  mutate(child_id = "Y") %>%
  left_join(meta)

if (saveFiles){
save(y_detections, file = paste0('../data/temp_data/Y_detections_preprocessed_',today(),'.RData'))
}
```

### Merge detections from Y with other kids
```{r}
# load(file='../data/temp_data/y_detections_preprocessed.RData')
y_detections <-  y_detections %>%
  rename(hand_openpose = hand_detected, face_openpose = face_detected) # standardize

d_all <- d_SA %>%
  mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
  full_join(y_detections)

# check
length(unique(d_all$vid_name)) == (length(unique(y_detections$vid_name)) +  length(unique(d_SA$vid_name)))

# clean up
remove(d_SA)
remove(d_y)

# save
if (saveFiles){
save(d_all, file = paste0('../data/preprocessed_data/openpose_detections_filtered_',today(),'.RData'))
}
```

### Look at any suspicious Sam videos with very high face detections (candidate missed allocentric videos)
```{r eval=FALSE}
# vid_check <- d_all %>%
#   filter(child_id == "S") %>%
#   group_by(vid_name, Databrary.Link) %>%
#   summarize(prop_faces=mean(face_openpose))
# 
# high_faces <- vid_check %>%
#   filter(prop_faces>.8)
```

# Cropped detections
## Load bounding boxes from S/A to get cropped detections
```{r}
## import bounding boxes
load('../data/raw_data/SA_bounding_boxes.RData')
bb_SA <- df
remove(df)

# get shortened video names so they match across all files
all_video_names_bb = as_tibble(unique(bb_SA$vid_name))
all_video_names_bb <- all_video_names_bb %>%
  rowwise() %>%
  mutate(vid_name = str_split(value,"[.]")[[1]][1])

# reset levels of vid_name in big detections array for detections & bb (very slow rowwise)
bb_SA$vid_name <- as.factor(bb_SA$vid_name)
levels(bb_SA$vid_name) <- all_video_names_bb$vid_name

# filter out flipped videos & get rid of bad variables
bb_SA <- bb_SA %>%
  filter(!vid_name %in% vids_to_exclude$vid_name) %>%
  select(-right_side_up, -age_days) # these variables were computed on the video names in other code and have NAs for some weird vids, excluding

# check that we excluded the wrong videos here 
vids_bb_S = length(unique(bb_SA$vid_name[bb_SA$child_id=='S']))
vids_bb_A = length(unique(bb_SA$vid_name[bb_SA$child_id=='A']))
vids_dets_S = length(unique(d_all$vid_name[d_all$child_id=='S']))
vids_dets_A = length(unique(d_all$vid_name[d_all$child_id=='A']))

## these are two videos in which A is just staring at the ceiling, no no bb's to be had in the entire vids
## can look at Databrary.link
vids_no_bb <- all_video_names %>%
  filter(!vid_name %in% all_video_names_bb$vid_name) %>%
  left_join(meta)

#
assert_that(vids_bb_S == vids_dets_S)

# if we meet checks, save
if (saveFiles){
save(bb_SA, file = paste0('../data/temp_data/SA_bounding_boxes_cleaned_', today(), '.RData'))
}
```

## Join detections for both and save out for later
```{r}
# load(file = '../data/Y_bounding_boxes_cleaned.RData') #  d_y from earlier if not loaded

## frame
height_px = 480; width_px = 640
bb_all_three <- d_y %>%
  filter(person==0) %>% # only rows with detections for now so comparable to bb_all
  # mutate(left = left*width_px, top = top*height_px, height = height*height_px, width = width*width_px) %>% #unnormalize data so comparable
  full_join(bb_SA) %>% # join with SA bounding boxes
  mutate(center_x = left + width/2, center_y = top - height/2, area = width*height)

save(bb_all_three, file = '../data/all_bounding_boxes_three_kids_preprocessed.RData')
remove(bb_all_three)
```

# Load temp files if we are skipping steps above run if we are skipping steps above
```{r eval=FALSE}
## For cropped detections, need..
# 1. full set of detections (d_all from above)
load(file='../data/preprocessed_data/openpose_detections_filtered_2020-01-29.RData')

# 2. bounding boxes from S and A (bb_SA)
load(file='../data/temp_data/SA_bounding_boxes_cleaned_2020-01-29.RData')

# 3. bounding boxes from Y (d_y) -- has different format than bb_SA, so diff name..
load(file='../data/temp_data/Y_bounding_boxes_cleaned_2020-01-29.RData')
```

### Make cropped detections for Y/SA, join, make cropped large data
```{r}
detections_cropped_Y <- d_y %>%
  filter(person>0) %>% #only get frames where detections so comparable
  mutate(top_real = top - height) %>% ## TOP IS ACTULALY BOTTOM AND 0,0 IS in the "TOP/LEFT" ! OMG.
  mutate(center_y = top_real + height/2, center_x = left + width/2) %>%
  filter(center_y<(.6*480)) %>%  ## 
  group_by(vid_name, frame, child_id, age_days) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))

detections_cropped_Y <- detections_cropped_Y %>%
 ungroup() %>%
 mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id))

## Make cropped detections in top 60% of frame for S,A
detections_cropped_SA <- bb_SA %>%
  mutate(top_real = top - height) %>% ## TOP IS ACTULALY BOTTOM AND 0,0 IS TOP
  mutate(center_y = top_real + height/2, center_x = left + width/2) %>%
  filter(center_y<(.6)) %>%  ## 
  group_by(vid_name, frame, child_id) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))

# join all together
all_cropped_detections <- detections_cropped_SA %>%
  ungroup() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  full_join(detections_cropped_Y)

## clean up workspace or else we run into memory limits
# cropped detections
remove(detections_cropped_Y)
remove(detections_cropped_SA)

# bb data
remove(y_data)
remove(bb_all)

## need to get "FALSE" on frames where we didn't have info to be able to compute proportions; uses d_all from earlier and fills it all back in

all_cropped_detections <- all_cropped_detections %>%
  select(-age_days, -child_id)

d_cropped <- d_all %>% ## needs to be loaded!
  as_tibble() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  select(vid_name, frame, child_id, age_days, face_openpose, hand_openpose) %>%
  left_join(all_cropped_detections, by=c("vid_name","frame"))

if (saveFiles){
save(d_cropped, file = paste0('../data/preprocessed_data/openpose_detections_filtered_cropped', today(), '.RData'))
}
```

### Make center detections for Y/SA, join, make centered large data
```{r}
limit_lower =.2
limit_upper =.8
y_max = 480
x_max = 640

detections_center_Y <- d_y %>%
  filter(person>0) %>% #only get frames where detections so comparable
  mutate(top_real = top - height) %>%
  mutate(center_y = top_real + height/2, center_x = left + width/2) %>%
  filter(center_y<(limit_upper*y_max) & center_x<(limit_upper*x_max) & center_y>(limit_lower*y_max) & center_x>(limit_lower*x_max)) %>%  ## 
  group_by(vid_name, frame, child_id, age_days) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))

detections_center_Y <- detections_center_Y %>%
 ungroup() %>%
 mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id))

## Make center detections in top 60% of frame for S,A
detections_center_SA <- bb_SA %>%
  mutate(top_real = top - height) %>%
  mutate(center_y = top_real + height/2, center_x = left + width/2) %>%
  filter(center_y<(limit_upper) & center_x<(limit_upper) & center_y>(limit_lower) & center_x>(limit_lower)) %>%  ## 
  group_by(vid_name, frame, child_id) %>%
  summarize(hand_detected = (sum(label=="hand")>0), face_detected = (sum(label=="face")>0), num_hands = sum(label=="hand"), num_faces = sum(label=="face"), num_people = sum(label=="pose"))  
# join all together
all_center_detections <- detections_center_SA %>%
  ungroup() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  full_join(detections_center_Y)
```

## Join back with big data structure
```{r}
## clean up workspace or else we run into memory limits
# center detections
remove(detections_center_Y)
remove(detections_center_SA)

# # bb data
# remove(y_data)
# remove(bb_all)

## need to get "FALSE" on frames where we didn't have info to be able to compute proportions; uses d_all from earlier and fills it all back in
all_center_detections <- all_center_detections %>%
  select(-age_days, -child_id) %>%
  mutate(face_openpose_center = face_detected, hand_openpose_center = hand_detected)

d_center <- d_all %>% ## needs to be loaded!
  as_tibble() %>%
  mutate(vid_name = as.factor(vid_name), child_id = as.factor(child_id)) %>%
  select(vid_name, frame, child_id, age_days, face_openpose, hand_openpose) %>%
  left_join(all_center_detections, by=c("vid_name","frame"))

if (saveFiles){
save(d_center, file = paste0('../data/preprocessed_data/openpose_detections_filtered_center2', today(), '.RData'))
}
```

```{r}
### Code to start to see if it maatters how many people are in the frame
age_info <- d_cropped %>%
  select(vid_name, frame, child_id, age_days)

one_person <- detections_cropped_SA %>%
  filter(num_people < 2) %>%
  left_join(d_cropped) %>%
  group_by(age_days, child_id) %>%
  summarize(prop_hands = mean(hand_detected), prop_faces = mean(face_detected))

wide_one_person <- one_person %>%
  gather(region, prop, prop_hands, prop_faces)

ggplot(wide_one_person, aes(x=age_days, y=prop, col=region)) +
  geom_point() +
  geom_smooth() +
  facet_grid(.~child_id) +
  ggtitle('hands vs faces (cropped) -- frames with only 1 person')
```

# Other data files that get cleaned up for cogsci
### Gold sample annotations
```{r}
## load goal sample annotations
load('../data/raw_data/ketan_gold_sample.RData')

## preprocess them
 ketan_gold_out <- ketan_gold %>%
   rowwise() %>%
   mutate(vid_name = str_split_fixed(vid_name,".mp4",2)[,1])%>%
   mutate(vid_name_short = str_split(vid_name,"[.]")[[1]][1]) %>%
   filter(!vid_name_short %in% flipped_videos$video) %>%
   mutate(face_openpose = as.logical(face_openpose), hand_openpose = as.logical(hand_openpose)) %>%
   rename(face_present_ketan = face_present, hand_present_ketan = hand_present)

gold_sample <- ketan_gold_out %>%
  select(-vid_name) %>%
  rename(vid_name = vid_name_short) %>% 
  mutate(vid_name = as.factor(vid_name)) %<%
  filter(!vid_name %in% to_exclude$vid_name)

if (saveFiles){
save(gold_sample, file = paste0('../data/preprocessed_data/gold_sample_annotations', today(), '.RData'))
}
```

### Turk annotations (get rid of identifying info)
```{r}
turk_annotations_seg = "../data/raw_data/turk_segmentations_hands_only_processed.csv" 
hand_annotations = read.csv(turk_annotations_seg) %>%
  select(-HITID, -HITTypeId, -WorkerID) %>%
  mutate(center_x = left + width/2, center_y = top + height/2)

if (saveFiles){
save(hand_annotations, file = paste0('../data/preprocessed_data/hand_annotations_', today(), '.RData'))
}

```

```{r}
# what <- d_cropped %>%
#   filter(face_openpose == TRUE) %>%
#   filter(fa)
# 
child_hands_cropping <- d_cropped %>%
  replace_na(list(hand_detected = FALSE, face_detected = FALSE)) %>%
  group_by(vid_name) %>%
  summarize(prop_faces = mean(face_openpose), prop_hands = mean(hand_openpose), prop_faces_cropped = mean(face_detected), prop_hands_cropped = mean(hand_detected), diff = prop_hands - prop_hands_cropped) %>%
  filter(diff>.3) %>%
  left_join(meta)

vid_summary <- vid_summary %>%
  left_join(meta)
# weird_vids = unique(vid_summary$vid_name[vid_summary$prop_hands_cropped==0])
# 
# weird_d_all <- d_all %>%
#   filter(vid_name %in% weird_vids)
# 
# bb_all_check <- bb_all %>%
#   filter(vid_name %in% weird_vids)
```