### 2021-04-16 analysis/bounding_boxes_april2021.Rmd
initial bounding box analyses script

---
title: "bounding_box_analyses_2021"
author: "Bria Long"
date: "4/15/2021"
output: html_document
---

```{r}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(stringr)
library(here)
theme_set(theme_few())
```

```{r}
load(file=here::here('data/preprocessed_data_2021/bounding_box_summaries.RData'))
```

## construct age bins for plotting
```{r}
bin_size=90
min_age = min(all_vid_bbs$age_days, na.rm=TRUE)
max_age = max(all_vid_bbs$age_days, na.rm=TRUE)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)

all_vid_bbs <- all_vid_bbs %>%
  mutate(age_months = age_days/30.4) %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) 
```

## restrict to face/hands for now
```{r}
face_hands <- all_vid_bbs %>%
  filter(!label=='pose') %>%
  mutate(avg_center_y = avg_top + avg_height/2, avg_center_x = avg_left + avg_width/2)
```

```{r}
ggplot(face_hands, aes(x=age_days/30.4, y=log(avg_area), col=label, size=num_detect)) +
  geom_point(alpha=.1) +
  facet_grid(~child_id) +
  geom_smooth(span=10000000) + # lol what
  ylab('Log (average area)') +
  xlab('Age in days')
```

```{r}
ggplot(face_hands, aes(x=age_days, y=log(avg_height), col=label, size=num_detect)) +
  geom_point(alpha=.1) +
  facet_grid(~child_id) +
  geom_smooth(span=10000000) + # lol what
  ylab('Log (average height)') +
  xlab('Age in days')
```

```{R}
ggplot(faces, aes(x=log(avg_area))) +
  geom_histogram(bin=100) +
  facet_grid(~child_id) +
  ggtitle('Log (average area) of faces') 
```

```{r}
ggplot(face_hands %>% filter(log(avg_area)>-10), aes(x=log(avg_area), col=label, fill=label)) +
  geom_histogram(alpha=.1, bins=100) +
  facet_wrap(~age_day_bin) +
  ggtitle('Average sizes of faces/hands') 
```

## Face hand density by age
```{r}
ggplot(face_hands, aes(x=avg_center_x , y = avg_center_y, size=num_detect, color=label)) +
  geom_point(alpha=.3) + 
  stat_density2d(aes(fill = label, color=label), geom="polygon", alpha=.6) +
  coord_fixed(ratio=.76) +
  ggtitle('Face/hand density') +
  ylab('') +
  xlab('') +
  theme_few(base_size=14) +
  theme(legend.position="right", axis.text.x=element_blank(), axis.ticks.x=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()) +
  scale_y_reverse() +
  facet_wrap(~age_day_bin)

```