### 2021-05-04 preprocessing/step5_preprocess_fullset_from_bb_outputs.Rmd

 ---
title: "SAYCAM preprocessing analyses from bounding box outputs"
author: "Bria Long"
date: "6/2020, updated 4/2021"
output:
  pdf_document: default
  html_document: default
---

# Basic data loading and setup
## Packages
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(stringr)
library(here)
theme_set(theme_few())
```

## Info to compute ages
```{r}
bday_s_approx = lubridate::ymd(20121017)
bday_a_approx = lubridate::ymd(20120913)
bday_y_approx = lubridate::ymd(20180214)

bday = c(bday_s_approx,bday_a_approx, bday_y_approx)
child_id  = c('S','A','Y')
age_info = tibble(bday,child_id)
```

## Paths to raw outputs & datafiles
```{r}
# toggle between each of these folders / dataset types
dataset_dir = "/data/op_outputs_per_vid_rds/"; dataset_type = "all_detections"
# dataset_dir = "/data/op_outputs_per_vid_rds_high_conf/"; dataset_type = "high_conf_detections"

op_file_dir = here::here(dataset_dir)
op_files = dir(op_file_dir)
# op_files = op_files [1] # for testing

allocentric_file = here::here("data/raw_data/SAYCAM_allocentric_videos.csv")
to_exclude =read.csv(allocentric_file)

# get INCORRECT videos where allocentric videopoint 
to_exclude <- to_exclude %>%
  filter(allocentric==1) %>%
  rename(vid_name = video)
```

## Main detections: read in bounding box outputs from each vid, summarize and compile.
```{r}
reload_data=TRUE

# cropping paramts
limit_lower =.2
limit_upper =.8
# vid params
y_max = 480
x_max = 640

if (reload_data){
vid_index = 0

for (vid_file in op_files) {
  vid_index = vid_index + 1

  vid_file_path = paste0(here::here(), dataset_dir, vid_file)
  this_vid_data <- readRDS(vid_file_path)
   #
  assert_that(this_vid_data$frame[1]==0) # make sure we have all frames in this rds

   # check vid name to see if excluded vid
   this_vid_name_temp <- this_vid_data$vid_name[1]
   this_vid_name <- str_split(this_vid_name_temp,"[.]")[[1]][1]

   # only run if not excluded
   if (!this_vid_name %in% to_exclude$vid_name){

   # replace negative eye distances with NAs
   this_vid_data <- this_vid_data %>%
      mutate(eye_distance = replace(eye_distance, eye_distance<0,NA))

   # make sure we only have high conf detectino
   if (dataset_type=='high_conf_detections'){
      this_vid_data <- this_vid_data %>%
         filter(conf_thres==.5)
      assert_that(length(this_vid_data$frame)>0)
   }

    # summarize hands/faces in each video
   summary_vid_data <- this_vid_data %>%
     group_by(frame, child_id, vid_name, full_face, eye_distance) %>%
     summarize(hand_detected = sum(!is.na(mean_confidence[label=='hand']))>0,face_detected = sum(!is.na(mean_confidence[label=='face']))>0)  %>%
     group_by(child_id, vid_name) %>%
     summarize(num_hands = sum(hand_detected), num_faces = sum(face_detected), num_full_faces = sum(full_face, na.rm=TRUE), avg_eye_distance = mean(eye_distance, na.rm=TRUE))

   # full_face_data <- this_vid_data %>%
   #    filter(label=='face') %>%
   #    group_by(full_face) %>%
   #    summarize(eye_dist = mean(eye_distance,na.rm=TRUE))

   # get face/hand detections that occured together
   faces_and_hands <- this_vid_data %>%
      group_by(frame, child_id, vid_name) %>% # group at frame level first
     summarize(hand_detected = sum(!is.na(mean_confidence[label=='hand']))>0,face_detected = sum(!is.na(mean_confidence[label=='face']))>0) %>%
      filter(hand_detected==TRUE) %>%
      group_by(child_id, vid_name) %>% # now at the video level
     summarize(faces_and_hands = sum(face_detected)) # count how many faces/hands

   # crop detections to center and get hands/faces
  this_vid_data_center <- this_vid_data %>%
     mutate(center_y = top + height/2, center_x = left + width/2) %>% # get center_x and y
     filter(center_y<(limit_upper) & center_x<(limit_upper) & center_y>(limit_lower) & center_x>(limit_lower)) %>%   # filter within parameters
     group_by(frame, child_id, vid_name) %>% # group at frame level first
     summarize(hand_detected_center = sum(!is.na(mean_confidence[label=='hand']))>0, face_detected_center = sum(!is.na(mean_confidence[label=='face']))>0)

   faces_and_hands_center <- this_vid_data_center %>%
      filter(hand_detected_center==TRUE) %>%
      group_by(child_id, vid_name) %>% # now at the video level
     summarize(face_and_hands_center = sum(face_detected_center)) # count how many faces/hands  together

    # join datasets for center cropped
    summary_vid_data_center <- this_vid_data_center %>%
       group_by(child_id, vid_name) %>% # now at the video level
       summarize(num_hands_center = sum(hand_detected_center), num_faces_center = sum(face_detected_center)) %>%  # count how many faces/hands
      left_join(faces_and_hands_center)

   #  save conf thres & num franes
   this_conf_thres = mean(this_vid_data$conf_thres,na.rm=TRUE)
   num_frames = length(unique(this_vid_data$frame)) # count frames before doing anything

  # create summary from both normal and cropped, clean vid_name, get age info
  joined_summary <- summary_vid_data %>%
    left_join(faces_and_hands) %>%
    left_join(summary_vid_data_center) %>%
    rowwise %>%
    mutate(vid_name = str_split(vid_name,"[.]")[[1]][1]) %>% 
    left_join(age_info) %>%
    mutate(date_filmed = lubridate::ymd(str_split_fixed(vid_name,"_",4)[[2]])) %>%
    mutate(age_days = as.numeric(difftime(date_filmed,bday)) ) %>%
    select(-bday, -date_filmed) %>%
    mutate(conf_thres = this_conf_thres) %>%
    mutate(num_frames = num_frames)

  # join it: 1 row = 1 video
   if (vid_index == 1 ){
     all_vid_data <- joined_summary 
   }
   else if (vid_index > 1) {
   all_vid_data <- all_vid_data %>%
     full_join(joined_summary)
   }

    # assert that we have actual center face/hands data 
  # check that length of videos without face/hands in center is not equal to the num of videos in dataframe, 
  assert_that(!sum(is.na(all_vid_data$face_and_hands_center))==length(all_vid_data$vid_name))
}
}
}
```

### Choose bins across which to aggregate data & save out
```{r}
bin_size = 7 # days
min_age = min(all_vid_data$age_days, na.rm=TRUE)
max_age = max(all_vid_data$age_days, na.rm=TRUE)
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)

all_vid_data <- all_vid_data %>%
  mutate(age_day_bin = cut(age_days, bins, labels=round(bin_starts/30,1))) %>%
  mutate(age_day_bin = as.numeric(as.character(age_day_bin))) 
```

```{r}
save(all_vid_data, file = here::here(paste0('data/preprocessed_data_2021/all_vid_data_from_bbs_', dataset_type, '.RData')))

```

## Gold sample detections only -- specific frame

### Get detections from gold sample videos only -- have to do this to get the SPECIFIC FRAMES (not videos) from which things were sampled....
```{r}
reload_data=TRUE
# redo gold sample 
if (reload_data){
vid_index = 0
## load gold sample annotations

load(here('data/data_cogsci/gold_sample_annotations2020-01-31.RData')) # gold_sample
gold_sample_frames <- gold_sample %>%
  select(-X, -level_0) %>%
  select(vid_name, frame) %>%
  mutate(vid_name = str_split(vid_name,"[.]")[[1]][1])

gold_sample_vids <- unique(gold_sample_frames$vid_name)

for (vid_file in op_files) {

   # to check progress
   vid_index = vid_index + 1
   if ((vid_index %% 100)==0) { print (vid_file) }

   # get vid level data
   vid_file_path = paste0(here::here(), dataset_dir, vid_file)
   this_vid_data <- readRDS(vid_file_path)
   this_vid_name_temp <- this_vid_data$vid_name[1]
   this_vid_name <- str_split(this_vid_name_temp,"[.]")[[1]][1]

   # if it's one of the vids from gold sample
   if (this_vid_name %in% gold_sample_vids){
   # and not excluded
   if (!this_vid_name %in% to_exclude$vid_name){

   frames_to_get <- gold_sample_frames %>%
     filter(vid_name == this_vid_name)

   # now get vid level data
   gold_sample_op <- this_vid_data %>%
     filter(frame %in% frames_to_get$frame) %>%
     mutate(vid_name = str_split(vid_name,"[.]")[[1]][1]) %>%
     group_by(frame, vid_name, child_id) %>%
     summarize(hand_detected = sum(!is.na(mean_confidence[label=='hand']))>0,face_detected = sum(!is.na(mean_confidence[label=='face']))>0)

   # compile vid-level summaries 
   if (vid_index == 1){
     all_gold_sample_frames_op <- gold_sample_op
   }
   else if (vid_index > 1) {
   all_gold_sample_frames_op <- all_gold_sample_frames_op %>%
     full_join(gold_sample_op)
   }
}
}
}
}

## Add in age data
all_gold_sample_frames_op <-all_gold_sample_frames_op %>%
  left_join(age_info) %>%
  mutate(date_filmed = lubridate::ymd(str_split_fixed(vid_name,"_",4)[[2]])) %>%
  mutate(age_days = as.numeric(difftime(date_filmed,bday)) )

save(all_gold_sample_frames_op, file = here::here('data/preprocessed_data_2021/',paste0('gold_sample_from_bbs_2021_',dataset_type,'.RData')))
```