### 2021-05-25 preprocessing/step4_create_bounding_box_summaries.Rmd
minor updates to bounding box summaries csript

---
title: "bounding_box_preprocesing"
author: "Bria Long"
date: "7/16/2020"
output: html_document
---

# Basic data loading and setup
## Packages
```{r}
knitr::opts_chunk$set(echo = TRUE)
library(knitr)
library(lubridate)
library(tidyverse)
library(assertthat)
library(ggthemes)
library(lme4)
library(langcog)
library(viridis)
library(stringr)
library(here)
theme_set(theme_few())
```

## Info to compute ages
```{r}
bday_s_approx = lubridate::ymd(20121017)
bday_a_approx = lubridate::ymd(20120913)
bday_y_approx = lubridate::ymd(20180214)

bday = c(bday_s_approx,bday_a_approx, bday_y_approx)
child_id  = c('S','A','Y')
age_info = tibble(bday,child_id)
```

## Paths to raw outputs & datafiles
```{r}
op_file_dir = here::here("data/op_outputs_per_vid_rds/")
op_files = dir(op_file_dir)
# op_files = op_files [1] # for testing

allocentric_file = here::here("data/raw_data/SAYCAM_allocentric_videos.csv")
to_exclude =read.csv(allocentric_file)

# get INCORRECT videos where allocentric videopoint 
to_exclude <- to_exclude %>%
  filter(allocentric==1) %>%
  rename(vid_name = video)
```

## Main detections: read in bounding box outputs from each vid, summarize and compile.
```{r}
reload_data=TRUE
# instead of loading each vid, load summary...

# vid params
y_max = 480
x_max = 640

if (reload_data){
vid_index = 0
for (vid_file in op_files) {
  vid_index = vid_index + 1
   vid_file_path = paste0(here::here(), '/data/op_outputs_per_vid_rds/', vid_file)
   this_vid_data <- readRDS(vid_file_path)

   # # if data wasn't normalized -- some (but not all!) Y videos
   # if (this_vid_data$child_id[1]=='Y' &  max(this_vid_data$height, na.rm=TRUE)>1) {
   #   this_vid_data <- this_vid_data %>% 
   #      mutate(top = top/y_max, height = height/y_max, left = left/x_max, width = width/x_max)
   # }

   # this_vid <- this_vid_data %>%
   #    filter(!is.na(height)) %>%
   #    filter(mean_confidence > 0) %>% # sanity
   #    mutate(area = height * width) %>%
   #    group_by(label, child_id, vid_name) %>%
   #    summarize(num_detect = length(height), avg_area = median(area), avg_height = median(height), avg_width = median(width), avg_left = median(left), avg_top = median(top))

   this_vid_faces <- this_vid_data %>%
      filter(!is.na(height)) %>%
      filter(label=='face') %>%
      filter(mean_confidence > 0)

   empty_vid=FALSE
   # empty_vid = (length(this_vid_data$num_detect)==0)
   # if (empty_vid) {
   #   print(vid_file)
   #  }

   if (vid_index == 1 ){
     # all_vid_bbs <- this_vid
     all_faces <- this_vid_faces
   }
   else if (vid_index > 1 && empty_vid==FALSE) {
   # all_vid_bbs <- all_vid_bbs %>%
   #   full_join(this_vid)

   all_faces <- all_faces %>%
      full_join(this_vid_faces)
   }
}
}
```

```{r}
# all_vid_bbs <- all_vid_bbs %>%
#   left_join(age_info) %>%
#   rowwise() %>%
#   mutate(date_filmed = lubridate::ymd(str_split_fixed(vid_name,"_",4)[[2]])) %>%
#   mutate(age_days = as.numeric(difftime(date_filmed,bday)) )
# 
# save(all_vid_bbs,file=here::here('data/preprocessed_data_2021/bounding_box_summaries_all.RData'))

all_faces <- all_faces %>%
  left_join(age_info) %>%
  rowwise() %>%
  mutate(date_filmed = lubridate::ymd(str_split_fixed(vid_name,"_",4)[[2]])) %>%
  mutate(age_days = as.numeric(difftime(date_filmed,bday)) )

 save(all_vid_bbs,file=here::here('data/preprocessed_data_2021/all_faces_summary.RData'))

```