### 2022-01-06 paper/manuscript/render_annotations.Rmd

---
title: "render_annotations"
author: "Bria Long"
date: "12/22/2021"
output: html_document
---

```{r setup, echo=FALSE}
library(tidyverse)
library(here)
library(ggthemes)
library(magick)
library(viridis)
```

```{r, echo=FALSE}
load(file = here::here('data/preprocessed_data_2021/goldset_category_annotations.RData'))
```

```{r, echo=FALSE}
load(file = here::here('data/preprocessed_data_2021/gold_sample_from_bbs_2021_all_detections.RData'))

```

```{r}
load(file=here::here('data/preprocessed_data_2021/all_vid_data_from_bbs_all_detections.RData')) # all detections
bin_size = 7 # days
min_age = min(all_vid_data$age_days, na.rm=TRUE) # 182
max_age = max(all_vid_data$age_days, na.rm=TRUE) # 966
bin_starts = seq(min_age-1, max_age+1,bin_size)
bins = c(bin_starts, max_age)

```

## Render out child hand examples
```{r}

  child_hands <- d %>%
    filter(child_hand_seg == TRUE)

  public_urls = unique(child_hands$public_url)
  img_names_short = unique(child_hands$img_name)

  dir.create(here::here("data/annotation_examples/child_hands_take2"),recursive = TRUE)

  for (i in 1:length(public_urls)){
    this_image = image_read(public_urls[i])
    image_write(this_image,here::here("data/annotation_examples/child_hands_take2", img_names_short[i]))
  }

```

```{r}
  child_hands <- d %>%
    filter(child_hand_seg == TRUE) %>%
    filter(category=='Food')

  public_urls = unique(child_hands$public_url)
  img_names_short = unique(child_hands$img_name)

  dir.create(here::here("data/annotation_examples/hands_food"),recursive = TRUE)

  for (i in 1:length(public_urls)){
    this_image = image_read(public_urls[i])
    image_write(this_image,here::here("data/annotation_examples/hands_food", img_names_short[i]))
  }
```

```{r}
  hands_toys <- d %>%
    filter(child_hand_seg == TRUE) %>%
    filter(category=='Toy')

  public_urls = unique(hands_toys$public_url)
  img_names_short = unique(hands_toys$img_name)

  dir.create(here::here("data/annotation_examples/hands_toys"),recursive = TRUE)

  for (i in 1:length(public_urls)){
    this_image = image_read(public_urls[i])
    image_write(this_image,here::here("data/annotation_examples/hands_toys", img_names_short[i]))
  }
```

```{r}
  hands_toys <- d %>%
    filter(child_hand_seg == TRUE) %>%
    filter(category=='Toy')

  public_urls = unique(hands_toys$public_url)
  img_names_short = unique(hands_toys$img_name)

  dir.create(here::here("data/annotation_examples/hands_toys"),recursive = TRUE)

  for (i in 1:length(public_urls)){
    this_image = image_read(public_urls[i])
    image_write(this_image,here::here("data/annotation_examples/hands_toys", img_names_short[i]))
  }
```

```{r}
  dishes <- d %>%
    filter(child_hand_seg == TRUE) %>%
    filter(category=='Utensil-Dish')

  public_urls = unique(dishes$public_url)
  img_names_short = unique(dishes$img_name)

  dir.create(here::here("data/annotation_examples/hands_dishes"),recursive = TRUE)

  for (i in 1:length(public_urls)){
    this_image = image_read(public_urls[i])
    image_write(this_image,here::here("data/annotation_examples/hands_dishes", img_names_short[i]))
  }
```

```{r}
ugh_books <- d %>%
  filter(basic_level == 'book') %>%
  group_by(public_url, img_name) %>%
  mutate(book_ok = (category=='Book')) %>%
  filter(book_ok == FALSE)

  public_urls = unique(ugh_books$public_url)
  img_names_short = unique(ugh_books$img_name)

  dir.create(here::here("data/images/annotation_examples/ugh_books"),recursive = TRUE)

  for (i in 1:length(public_urls)){
    this_image = image_read(public_urls[i])
    image_write(this_image,here::here("data/images/annotation_examples/ugh_books", img_names_short[i]))
  }
```

## Render out individual categories
```{r}
reallyRun =0
if (reallyRun){
class_list  = unique(d$category)

for (c in class_list) {
  examples <- d %>%
    filter(category == c) %>%
    dplyr::sample_n(200)

  public_urls = unique(examples$public_url)
  img_names_short = unique(examples$img_name)

  dir.create(here::here("data/annotation_examples/",c),recursive = TRUE)

  for (i in 1:length(public_urls)){
    this_image = image_read(public_urls[i])
    image_write(this_image,here::here("data/annotation_examples/", c, img_names_short[i]))
  }
}
}
```