### 2022-01-07 analysis/motor_milestones.Rmd
walking milestone analysis

---
title: "Motor Milestone Analysis"
output: pdf_document
---

```{r setup, include=FALSE, echo=F, message=F}
knitr::opts_chunk$set(echo = F, warning=F, message=F)
library(tidyverse)
library(lme4)
library(stargazer)
library(lmerTest)
library(ggeffects)
```

We examine whether the SAYcam participants' age at which they started walking significantly influenced the proportion of hands or faces that they saw, using mixed-effects regressions with per-child random intercepts.
We tried including random slopes by age and milestone, but the models did not converge and/or were singular. The R syntax for the linear regression was thus: prop_faces ~ age * walk + (1 | child_id).
We also tested inclusion of a quadratic term for age (prop_faces ~ poly(age,2) * walk + (1 | child_id)).
Children's age of achieving the milestone was dummy-coded (i.e., 0=unable to walk; 1=able to walk).

```{r}
## Load in gold sample detections
load(here::here('data/preprocessed_data_2022/gold_sample_annotations2020-01-31.RData')) # human annotations
load(here::here('data/preprocessed_data_2022/gold_sample_from_bbs_2021_with_centerall_detections.RData')) 
load(here::here('data/preprocessed_data_2022/all_vid_data_from_bbs_all_detections.RData')) # all detections

```

```{r}
## summary for all detections
face_hand_by_age <- all_vid_data %>%
  ungroup() %>%
  tidyr::replace_na(list(faces_and_hands=0)) %>%
  group_by(age_day_bin, child_id) %>%
  summarize(num_frames_total = sum(num_frames), 
            # prop faces overall
            num_faces = sum(num_faces), 
            num_hands = sum(num_hands), 
            prop_faces = num_faces / num_frames_total, 
            prop_hands = num_hands / num_frames_total,
            # in center FOV
            num_faces_center = sum(num_faces_center, na.rm=TRUE), 
            num_hands_center = sum(num_hands_center, na.rm=TRUE), 
            prop_faces_center = num_faces_center / num_frames_total, 
            prop_hands_center = num_hands_center / num_frames_total,
            # detailed face info
            prop_full_faces = sum(num_full_faces, na.rm=TRUE)/num_frames_total,
            prop_faces_and_hands = sum(faces_and_hands)/num_frames_total,
            # face/hand contingency
            prop_faces_with_hands = sum(faces_and_hands, na.rm=TRUE)/num_faces, 
            prop_hands_with_faces = sum(faces_and_hands, na.rm=TRUE)/num_hands 
            )

all_detections <- face_hand_by_age %>%
  filter(num_frames_total > 2000) %>% # eliminate small data point that skews scaling
  select(prop_faces, prop_hands, num_frames_total, age_day_bin, child_id) %>%
  pivot_longer(cols = c(prop_faces, prop_hands), names_to = "region", values_to = "prop") %>%
  mutate(approach = "uncropped",
         region = ifelse(region == "prop_faces","Faces","Hands"))

```

```{r, get-milestones, echo=F}
motor <- read_csv(here::here("data","raw_data", "saycam_walking_asq.csv"))
# A and S are sitting at 6mos
face_hand_by_age_milestones <- face_hand_by_age %>%
  mutate(sit = ifelse(child_id=="S" | child_id=="A", 1, 
                      ifelse(child_id=="Y" & age_day_bin > 8, 1, 0))) %>%
  mutate(cruise = ifelse((child_id=="S" & age_day_bin >= 9) | 
                      (child_id=="A" & age_day_bin >= 10) | 
                      (child_id=="Y" & age_day_bin > 9), 1, 0)) %>%
  mutate(walk = ifelse((child_id=="S" & age_day_bin > 10) |
                       (child_id=="A" & age_day_bin > 13) |
                       (child_id=="Y" & age_day_bin > 16), 1, 0)) %>%
  mutate(age = age_day_bin)

library(lmerTest)
# https://rpsychologist.com/r-guide-longitudinal-lme-lmer
# with random slopes and intercepts ((age_day_bin | child_id): is singular / not converging
# so we drop random slopes

# cruise
face_cruise <- lmer(prop_faces ~ age * cruise + (1 | child_id), 
     data=face_hand_by_age_milestones)
hand_cruise <- lmer(prop_hands ~ age * cruise + (1 | child_id), 
     data=face_hand_by_age_milestones)

face_cruise2 <- lmer(prop_faces ~ poly(age,2) * cruise + (1 | child_id), 
     data=face_hand_by_age_milestones)
hand_cruise2 <- lmer(prop_hands ~ poly(age,2) * cruise + (1 | child_id), 
     data=face_hand_by_age_milestones)

# walk
face_walk <- lmer(prop_faces ~ age * walk + (1 | child_id), 
     data=face_hand_by_age_milestones)
hand_walk <- lmer(prop_hands ~ age * walk + (1 | child_id), 
     data=face_hand_by_age_milestones)

face_walk2 <- lmer(prop_faces ~ poly(age,2) * walk + (1 | child_id), 
     data=face_hand_by_age_milestones)
hand_walk2 <- lmer(prop_hands ~ poly(age,2) * walk + (1 | child_id), 
     data=face_hand_by_age_milestones)

#summary(face_walk2)
#summary(hand_walk2)

# make models compatible with stargazer
class(face_cruise) <- "lmerMod"
class(hand_cruise) <- "lmerMod"
class(face_cruise2) <- "lmerMod"
class(hand_cruise2) <- "lmerMod"

class(face_walk) <- "lmerMod"
class(hand_walk) <- "lmerMod"
class(face_walk2) <- "lmerMod"
class(hand_walk2) <- "lmerMod"
```

<!--
## Cruising

#The quadratic term was justified for the hands model ($\chi^2(2)=16.29$, $p<.001$), but not for the faces model ($\chi^2(2)=5.77$, $p=.06$).
#As children became able to cruise, they saw significantly more faces ($\beta = 0.27$), but there was no significant effect on hands. 
#For faces, there was a significant negative interaction of age and cruising ($\beta = -0.03$, $p=0.025$), and a marginal positive effect of age ($\beta = 0.03$, $p=0.055$).
-->

```{r, include=F}
anova(face_cruise, face_cruise2) # marginal
anova(hand_cruise, hand_cruise2) # sig
```

```{r, results='asis', include=F}
stargazer(face_cruise, hand_cruise2, title="Change in face and hand prevalence by ability to cruise", 
          header=F, dep.var.labels = c("Proportion of Faces","Proportion of Hands"), nobs=F)
```

```{r, fig.width=6.5, fig.height=3, include=F}
f1 <- ggpredict(face_cruise, c("age", "cruise")) %>% plot(add.data=T) +
  xlim(5, 30) + ylim(0,1) + ggtitle("Predicted effect on faces seen") + 
  xlab("Age (months)") + ylab("Proportion of faces")

h1 <- ggpredict(hand_cruise2, c("age", "cruise")) %>% plot(add.data=T) +
  xlim(5, 30) + ylim(0,1) + ggtitle("Predicted effect on hands seen") + 
  xlab("Age (months)") + ylab("Proportion of hands")

ggpubr::ggarrange(f1, h1, nrow=1, common.legend=T)
```

<!-- ## Walking -->

The quadratic term was justified for both the hands model ($\chi^2(2)=22.15$, $p<.001$) and the faces model ($\chi^2(2)=10.21$, $p<.01$).
As children became able to walk, they saw significantly more faces ($\beta = 0.26$, $p=0.02$) and hands ($\beta = 0.32$, $p=0.01$). 
The linear and quadratic age terms and their interactions with walking were also significant (see Table 1).
The models' predicted conditional effects are shown in Figure 1.

```{r, include=F}
anova(face_walk, face_walk2) # sig
anova(hand_walk, hand_walk2) # sig
```

```{r, results='asis'}
stargazer(face_walk2, hand_walk2, title="Change in face and hand prevalence by ability to walk", 
          header=F, dep.var.labels = c("Proportion of Faces","Proportion of Hands"), nobs=F)
```

```{r, fig.width=6.5, fig.height=3}
f1 <- ggpredict(face_walk2, c("age", "walk")) %>% plot(add.data=T) +
  xlim(5, 30) + ylim(0,1) + ggtitle("Predicted effect on faces seen") + 
  xlab("Age (months)") + ylab("Proportion of faces")

h1 <- ggpredict(hand_walk2, c("age", "walk")) %>% plot(add.data=T) +
  xlim(5, 30) + ylim(0,1) + ggtitle("Predicted effect on hands seen") + 
  xlab("Age (months)") + ylab("Proportion of hands")

ggpubr::ggarrange(f1, h1, nrow=1, common.legend=T)
```